---
title: "mHLEA_mixedeffects"
author: "Carolina Guidolin"
date: "2023-11-24"
output: html_document
---

```{r}
library(lme4)
library(Matrix)
library(tidyverse)
library(LightLogR)
library(sjstats)
```

## Code for doing a mixed effect model 
Loading in the subjective light exposure data (mH-LEA) from all participants using LightLogR to incorporate it as State
```{r}
mHLEA_mixedeffects <- data.LL.binned %>%
  interval2state(mhlea_df) %>%
  select(-lightsource, file.name) 

##We only want to keep the first light exposure category indicated by the participants 
mHLEA_mixedeffects$State <- gsub("\\+(.+)", "", mHLEA_mixedeffects$State)

##Account for participants who reported "W+X" which means that they would be sleeping, so we need to substitute to X
mHLEA_mixedeffects$State <- gsub("W", "X", mHLEA_mixedeffects$State)


```

## Organising our data structure
We need a column containing hourly self-reported light exposure categories (categorical variable, our fixed effect) and a column containing the dependent variable median objective light exposure (mEDI). We also want our grouping variable to be the participant. 

```{r pressure, echo=FALSE}
mHLEA_mixedeffects <- mHLEA_mixedeffects %>%
  mutate(hour= hour(Datetime)) %>% #extract the hour information
           group_by(Date = as.Date(Datetime, tz = tz), hour, State, Id) %>% #for each Id, hour and State, calculate the median medi
           summarise(median_medi = median(MEDI, na.rm = TRUE)) %>%
           filter(!is.na(State), !is.na(median_medi))

#We tranform mEDI data into log scale
small_nonzero_value <- 0.0001

mHLEA_mixedeffects$median_medi[mHLEA_mixedeffects$median_medi == 0] <- small_nonzero_value

mHLEA_mixedeffects$log_median_medi <- log(mHLEA_mixedeffects$median_medi, base = 10)

mHLEA_mixedeffects$Id <- as.factor(mHLEA_mixedeffects$Id) #Id has to be categorical variable
#mHLEA_mixedeffects$hour <- as.ordered(mHLEA_mixedeffects$hour)
```

##Testing different mixed models
```{r}
#We start with the most complex model: random intercept + random slope
model1 <- lmer(log_median_medi ~ State + (1 + State | Id), REML = FALSE, data = mHLEA_mixedeffects)
summary(model1)
isSingular(model1)

qqnorm(resid(model))
qqline(resid(model))

##leads to singularity and converge warning. 

##So we decide to eliminate the random slope
model_norandomslope <- lmer(log_median_medi ~ State + (1|Id), REML = FALSE, data = mHLEA_mixedeffects, na.action= na.exclude)
summary(model_norandomslope)

plot(model_norandomslope)


isSingular(model_norandomslope) #FALSE - model not singular
coef(model_norandomslope)

ci_bootstrap <- confint(model_norandomslope, method="boot") #calculate the CI using bootstrap method


```

```{r}
#To evaluate our random intercept model, we need a null model. Here, we choose one with pure random effects
null_model = lmer(log_median_medi~(1| Id), REML = FALSE, data=mHLEA_mixedeffects,na.action= na.exclude)

summary(null_model)
anova(null_model, model_norandomslope) #compare our model against null model



```

```{r}
residuals <- residuals(model)

# Plotting histogram of residuals
hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
```

