---
title: "mlit_calculation"
author: "Carolina Guidolin"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This code calculates the light metric MLiT250
MLiT is a light metric that stands for Mean light timing above a certain threshold X. This was defined by Reid et al. (2014) as the average clock time of all data points above a certain X lux. Here, we want to set the threshold to 250 lux, because this is the daytime light exposure level which is recommended by Brown et al. (2022).

While pyLight offers all tools to calculate this and related metrics in Python, there is also some code available on GitHub to do this in R. 

First, let's import the packages which we need

```{r cars}
library(tidyverse)
library(readr)
library(dplyr)
library(hms)
library(lubridate)
library(LightLogR)
library(magrittr)
library(stringr)
library(tibble)
library(tidyr)
library(zoo)
library(cowplot)
library(scales)
library(grid)
library(gridExtra)
library(svglite)
library(lightdosimetry)
```

Now, we install and load the lightdosimetry package

```{r}
#remotes::install_github("tscnlab/LightLogR")

remotes::install_github("steffenhartmeyer/spectrace@develop")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Let's try this out for all participants at once
We will import the data using the LightLogR package
```{r}
path <- "D:/cyepi/code/chronotype_analysis"
tz <- "Europe/Berlin"

# Get a list of file paths in the specified directory
files <- list.files(path, pattern = ".txt", full.names = TRUE)

# Create an empty dataframe to store the combined data
#dataset.LL <- data.frame()

# Iterate over each file and import data
 #for (file in files) {
 # dataset <- import.ActLumus(file, tz = tz, auto.id = "^(\\d{3})")
  #dataset.LL <- rbind(dataset.LL, dataset)  # Bind the dataset to the combined dataframe
#}
```


```{r}
#Using LightLogR directly - not working
 dataset.LL <- import.ActLumus(files, tz = tz, auto.id = "^(\\d{3})")
```
First, we want to modify the datetime to only contain time information. This is because if we include date information, the MLiT calculation will also take into account the date.
```{r}
dataset.LL.mlit <- dataset.LL.all_filtered %>%
  mutate(Time = as.POSIXct(as_hms(Datetime), tz = tz))
```

## Now we calculate MLiT250
```{r}
timing <- 
  dataset.LL.mlit %>%
  group_by(Id) %>%
   summarise(
    mlit(MEDI, Time, 250),
    flit(MEDI, Time, 250),
    llit(MEDI, Time, 250)
  )
```

```{r}
mlit250 <- timing %>%
  rename("id" = Id) %>%
  select(id, mlit.250)
```

##Merge mlit and msf_sc dataframes
```{r}
merged_mlit_msf <- merge(mlit250, msf_sc, by = "id")
```


## Visualise the data
```{r plot}
#install.packages("ggpubr")
library(ggpubr)
merged_mlit_msf$color_gradient <- scales::rescale(merged_mlit_msf$msf_sc_rounded)
y_limits <- as.POSIXct(c("1970-01-01 12:30", "1970-01-01 18:00"), tz = tz)

scatter <- ggplot(merged_mlit_msf, aes(x = msf_sc_rounded, y =mlit.250,  color = color_gradient)) + 
  geom_jitter(width = 0.2, height = 0.2) + 
 ggtitle("Correlation between chronotype \nand timing of mEDI above 250 lx") + 
  xlab(bquote("Chronotype "(MSF[sc]))) + 
  ylab(bquote("Mean light (mEDI) timing above 250 lx  "(MLiT^250))) +
  geom_smooth(method='lm', formula= y~x, colour = "black", linewidth = 0.5, alpha = .3) + 
  stat_cor(method = "spearman") +
  scale_color_gradient(low = "#FFCC00", high = "#FF6633") +
  scale_y_datetime(limits = y_limits, breaks = date_breaks("1 hour"), labels = date_format("%H:%M"), timezone = tz) +
  theme_bw() +
  theme(aspect.ratio = 1) +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold")) + 
  theme(legend.position = "none") +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_blank())

  
```
## We want to try run a correlation

```{r}
# Convert the variables to numeric
#merged_mlit_msf$mlit.250 <- as.numeric(merged_mlit_msf$mlit.250)
#merged_mlit_msf$msf_sc_rounded <- as.numeric(merged_mlit_msf$msf_sc_rounded)

# Run the correlation test
#corr <- cor.test(
 # x = merged_mlit_msf$mlit.250,
 # y = merged_mlit_msf$msf_sc_rounded,
 # method = 'spearman'
#)

```
It seems like, although there is a trend towards a correlation between chronotype and mean light exposure above 250 lux, the coefficient indicates that it is quite weak (~0.4). 

## Try a different visualisation 
We want a boxplot 
```{r}
# Calculate quantiles
quantiles <- quantile(merged_mlit_msf$msf_sc_rounded, probs = c(0.1, 0.9))

# Create a new column 'chronotype_cat' based on the quantiles
merged_mlit_msf$chronotype_cat <- ifelse(
  merged_mlit_msf$msf_sc_rounded < quantiles[1], "early",
  ifelse(merged_mlit_msf$msf_sc_rounded >= quantiles[2], "late", "intermediate")
)

```

```{r}
violin <- ggplot(merged_mlit_msf, aes(x = chronotype_cat, y = mlit.250)) +
  geom_violin(alpha = 0.3, aes(fill=chronotype_cat), trim = FALSE) +
  geom_boxplot(width = 0.2)+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5) +
  scale_fill_manual(values = c("#FFCC00", "#FF9900", "#FF6633")) +
   scale_y_datetime(limits = y_limits, breaks = date_breaks("1 hour"), labels = date_format("%H:%M"), timezone = tz) +
  theme(legend.position = "none") +
#  xlab(bquote("Chronotype "(MSF[sc]))) + 
#  ylab(bquote("Mean light (mEDI) timing above 250 lx  "(MLiT^250))) +
  ggtitle("Timing of mEDI above 250 lx \nbased on chronotype") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold")) +
    theme(aspect.ratio = 1) +
  theme(legend.position = "none") +  
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_blank())
 

 
```
## Put the two graphs together 
```{r}
grid <- plot_grid(scatter, violin, labels = c("A", "B"),
          align = "v",
          nrow =1,
          scale = 1)

#create common x and y labels

y.grob <- textGrob(bquote("Mean light (mEDI) timing above 250 lx  "(MLiT^250)), 
                   gp=gpar(fontface="bold", col="black", fontsize=11), rot=90)

x.grob <- textGrob(bquote("Chronotype "(MSF[sc])), 
                   gp=gpar(fontface="bold", col="black", fontsize=11), just = "bottom")


 chrono_multiplot <- grid.arrange(arrangeGrob(grid, left = y.grob, bottom = x.grob))

```
## Save the plot 
```{r}
library(svglite)
ggsave("chronotype_multiplot_final.png", plot =  chrono_multiplot, width = 8, height = 4)
```


```{r}
#checking for timezones
merged_mlit_msf$msf_sc_rounded <- with_tz(merged_mlit_msf$msf_sc_rounded, tz = "Europe/Berlin")
merged_mlit_msf$mlit.250 <- with_tz(merged_mlit_msf$mlit.250, tz = "Europe/Berlin")
```

