---
title: "pre-processing"
author: "Carolina Guidolin"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data pre-processing step 1: detection of non-wear time

## Description of wear and non- wear time

The first thing we'd like to do is to provide some summary statistics to describe our particiants' wear and non wear time. We will start by describing the information obtained from the wear log and compare this within- and between- people.First, we start by analysis the wear log data from participant 212.

#### Loading packages 


```{r}
library(LightLogR)
library(tidyverse)
library(gt)
library(patchwork)
library(ggplot2)
library(ggpubr)
library(tidyr)
library(hms)

```

#### Loading actlumus data 
First, we need to import the ActLumus file where the button press information is contained. We will need this later.
```{r}
path <- "D:/cyepi/code/pre-processing/nonwear_detection/actlumus"
tz <- "Europe/Berlin"

# Get a list of file paths in the specified directory
actlumusfiles <- list.files(path, pattern = ".txt", full.names = TRUE)

# Create an empty dataframe to store the combined data
dataset.LL.all <- data.frame()

# Iterate over each file and import data
dataset.LL.all <- import$ActLumus(actlumusfiles, tz = tz, auto.id = "^(\\d{3})")

```
#### Every participant started at different times. We will have to specify start and end time manually
```{r}
filtered_time <- data.frame(
  Id = c("201", "202", "204", "205", "206", "208", "209", "210", "212", "213", "214", "215", "216", "218", "219", "221", "222", "223", "224", "225", "226", "227", "228", "229", "230", "231"),
  start = as.POSIXct(c("2023-08-14 12:40:00", "2023-08-14 13:26:00", "2023-08-14 14:43:00", "2023-08-28 15:42:00", "2023-08-28 14:16:00", "2023-09-04 12:36:00", "2023-09-04 14:01:00", "2023-09-04 15:20:00", "2023-09-11 13:03:00", "2023-09-11 15:06:00", "2023-09-25 14:56:00", "2023-10-02 11:30:00", "2023-10-02 14:31:00", "2023-10-16 11:18:00", "2023-10-16 12:40:00", "2023-10-23 11:22:00", "2023-10-23 13:10:00", "2023-10-30 11:34:00", "2023-10-30 13:24:00", "2023-10-30 15:05:00", "2023-10-30 16:34:00", "2023-10-30 17:44:00", "2023-11-06 11:02", "2023-11-06 12:34", "2023-11-06 14:02", "2023-11-06 17:12" )),    # Add start times
  end = as.POSIXct(c("2023-08-20 23:59:59", "2023-08-20 23:59:59", "2023-08-20 23:59:59", "2023-09-03 23:59:59", "2023-09-03 23:59:59", "2023-09-10 23:59:59", "2023-09-10 23:59:59", "2023-09-10 23:59:59", "2023-09-17 23:59:59", "2023-09-17 23:59:59", "2023-10-01 23:59:59", "2023-10-08 23:59:59", "2023-10-08 23:59:59", "2023-10-22 23:59:59" , "2023-10-22 23:59:59", "2023-10-29 23:59:59", "2023-10-29 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59"))            # Add end times
)

dataset.LL.all <- dataset.LL.all %>%
  inner_join(filtered_time, by = "Id") %>%
  filter(Datetime >= start, Datetime <= end) %>%
  select(-start, -end)

```

#### Loading in the wearlog data 
```{r}

files <- setwd("D:/cyepi/code/pre-processing/nonwear_detection/wearlog")
# Get the files names
wearlogfiles = list.files(pattern="*.csv")
# First apply read.csv, then rbind
wearlog_df = do.call(rbind, lapply(wearlogfiles, function(x) read.csv(x, stringsAsFactors = FALSE, sep = ";")))
```

### Frequency of reporting taking the light glasses off, putting them on, and taking them off before sleep
How often did the participants report taking off the light glasses? We know that wearlog_event = 0 and wearlog_past = 0 correspond to an "off" event (current and retrospectie), while wearlog_1 and wearlog_past = 1 correspond to a "on" event (current and retropsective). wearlog_event = 2 and wearlog_past = 2 correspond to taking he light glasses off to go to sleep. 

### Frequency of using the black bag for reporting light glasses off
When participants take off the light glasses (wearlog_event and wearlog_past = 0), they are instructed to place in in a black bag. If they forget to the bag, they can also report this in the app. So how often did they use the black bag upon removing the light glasses?

```{r}
wearlog_events <- wearlog_df %>%
  group_by(record_id) %>%
  summarise(
            wearlog_off_current = sum(wearlog_event == 0, na.rm = TRUE),
            wearlog_off_retrospective = sum(wearlog_past == 0, na.rm = TRUE),
          #  wearlog_on_total = sum(wearlog_event == 1 | wearlog_past == 1, na.rm = TRUE),
            wearlog_on_current = sum(wearlog_event == 1, na.rm = TRUE),
            wearlog_on_retrospective = sum(wearlog_past == 1, na.rm = TRUE),
          #  wearlog_sleep_total = sum(wearlog_event == 2 | wearlog_past == 2, na.rm = TRUE),
            wearlog_sleep_current = sum(wearlog_event == 2, na.rm = TRUE),
            wearlog_sleep_retrospective = sum(wearlog_past == 2, na.rm = TRUE),
            wearlog_anyevent_z1 = n(),
            wearlog_bagused_z2 = sum(wearlog_bag == 1 | wearlog_past_bag == 1, na.rm = TRUE),
            wearlog_bagnotused_z3 = sum(wearlog_bag == 0 | wearlog_past_bag == 0, na.rm = TRUE),
            wearlog_leavingtubingen_z4 = sum(wearlog_event == 3 | wearlog_past == 3, na.rm = TRUE),
            wearlog_reenteringtubingen_z5 = sum(wearlog_event == 4 | wearlog_past == 4, na.rm = TRUE))

```

### Frequency of using button press for reporting light glasses on, off, and off before sleep
Participants are also instructed to press the button every time they take off the light glasses as well as any time they put them back on. This is saved in the device file itself as "event", where event = 1 means a button press took place. 

For this, we need to combine the dataset.LL.all and the nonwear_df datasets

We identify numbers of button
```{r}
button_press <- dataset.LL.all %>%
  group_by(Id) %>%
  summarise(wearlog_buttonpress_z6 = sum(EVENT == 1)) %>%
  rename(record_id = Id)
```

Let's incorporate the button press information into the nonwear_df
```{r}
nonwear_all <- merge(wearlog_events, button_press, by = "record_id")

nonwear_all_long <- pivot_longer(nonwear_all, cols = c("wearlog_anyevent_z1", "wearlog_off_current","wearlog_off_retrospective", "wearlog_on_current", "wearlog_on_retrospective",  "wearlog_sleep_current", "wearlog_sleep_retrospective", "wearlog_bagused_z2", "wearlog_bagnotused_z3", "wearlog_buttonpress_z6", "wearlog_leavingtubingen_z4", "wearlog_reenteringtubingen_z5"), names_to = "event", values_to = "count")
```

```{r}
##Getting some summary statistics

nonwear_all_summary <- colSums(nonwear_all, na.rm = TRUE)
sums_df <- data.frame(
  Variable = names(nonwear_all_summary),
  Sum = column_sums
)

nonwear_all_summary <- nonwear_all %>%
  summarise(
    across(-record_id, list(Sum = sum), na.rm = TRUE),
    across(-record_id, list(ParticipantWithMax = ~record_id[which.max(.)])),
    across(-record_id, list(ParticipantWithMin = ~record_id[which.min(.)]))
  )

summary_table <- nonwear_all_long %>%
  group_by(event) %>%
  summarise(median = median(count), iqr = IQR(count)) %>%
  rename(Event = event)

summary_table$Event[summary_table$Event %in% c("wearlog_anyevent_z1", "wearlog_off_current","wearlog_off_retrospective", "wearlog_on_current", "wearlog_on_retrospective",  "wearlog_sleep_current", "wearlog_sleep_retrospective", "wearlog_bagused_z2", "wearlog_bagnotused_z3", "wearlog_buttonpress_z6", "wearlog_leavingtubingen_z4", "wearlog_reenteringtubingen_z5")] <- c("All wear log entries", "Taking the light glasses off (current)", "Taking the light glasses off (retrospective)", "Putting the light glasses on (current)", "Putting the light glasses on (retrospective)", "Taking the light glasses off before sleep (current)" ,"Taking the light glasses off before sleep (retrospective)","Black bag used", "Black bag not used", "Button press", "Leaving Tübingen", "Re-entering Tübingen")

summary_table <- gt(summary_table)

summary_table <- summary_table %>%
  tab_header(title = "Summary of self-reported \"Wear log\" entries and button presses for n=26 participants")     

summary_table
gt_preview(summary_table)

gtsave(summary_table, filename = "wearlog_table.png")
```



#### Let's visualise this information 
First, we start by visualising ALL events for all participants with a stacked barplot

We need to convert in long format for this:
```{r}
stacked_overview_long <- nonwear_all %>%
  pivot_longer(
    cols = -record_id,
    names_to = c("event", "time"),
    names_pattern = "wearlog_(\\w+)_(\\w+)",
    values_to = "count"
  )

summary_long <- stacked_overview_long %>%
  group_by(event, time) %>%
  summarise( median = median(count),
          lower_iqr = quantile(count, 0.25),
          upper_iqr = quantile(count, 0.75))
```

```{r}
custom_order_stacked <- c("reenteringtubingen", "leavingtubingen", "bagnotused", "bagused", "sleep", "off", "on", "buttonpress", "anyevent")

summary_plot_all <- ggplot(summary_long, aes(x = median, y = factor(event, custom_order_stacked), fill = time)) +
  geom_bar(position = "stack", stat = "identity")  +
  ggtitle(bquote("Number of total self-reported \"Wear log\" entries and\n light logger button press events for n=26 participants")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "plain")) +
  xlab("Count") +
  ylab("Event") +
  scale_y_discrete(breaks = custom_order_stacked, labels = c("Re-entering Tübingen", "Leaving Tübingen", "Off and black bag not used", "Off and black bag used", "Off before sleep ", "Off", "On", "Button press", "All wear log entries")) +
  scale_x_continuous(breaks = c(0,5,10,15,20,25,30,35), labels = c("0","5","10","15","20","25","30", "35"), expand = c(0,1)) +
scale_fill_manual(name = "Time of reporting", values = c( "#66CCFF", "#0066CC",  "#006633", "#CC66FF", "#FF99FF", "#FF9933", "#CC0033" , "#009933"), labels = c( "Current", "Retrospective", "z1", "z2", "z3","z4", "z5", "z6" )) +
  geom_errorbarh(
    aes(xmin = lower_iqr, xmax = upper_iqr),
    height = 0.3,
    color = "black",
    size = 0.5
  ) +
  theme(axis.title = element_text(size= 14),
        axis.text = element_text(size=12),
        plot.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size=12),
        legend.key.size = unit(0.3, "cm"),
        legend.position = "top",
        legend.direction = "horizontal",
        aspect.ratio = 0.9) 
```
Let's save this
```{r}
ggsave(plot = summary_plot_all, filename ="thesis_summary_events.png", dpi =600, width = 9, height = 6, path ="D:/cyepi/code/outputs")
ggsave(plot = summary_plot_all, filename ="thesis_summary_events.pdf", dpi =600, width = 9, height = 6, path ="D:/cyepi/code/outputs")
```

Another visualisation is to look at each individual participant in a facet
```{r}
faceted_plot_wearlog2 <- ggplot(stacked_overview_long, aes(x = count, y = factor(event, custom_order_stacked), fill = time)) +
  geom_bar(position = "stack", stat = "identity")  +
  facet_wrap(~record_id, ncol = 4, scales = "free_x") + 
  scale_y_discrete(breaks = custom_order_stacked, labels = c("Re-entering Tübingen", "Leaving Tübingen", "Off and black bag not used", "Off and black bag used", "Off before sleep ", "Off", "On", "Button press", "All wear log entries")) +
scale_fill_manual(name = "Time of reporting", values = c("#66CCFF", "#0066CC","#006633", "#CC66FF", "#FF99FF", "#FF9933", "#CC0033", "#009933"), labels = c( "Current", "Retrospective", "z1", "z2", "z3","z4", "z5", "z6" )) +
   ggtitle(bquote("Number of total self-reported \"Wear log\" entries and\n light logger button press events for n=26 participants")) + 
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "plain")) +
  xlab("Count") +
  ylab("Event") +
   theme_bw() +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top",
        legend.direction = "horizontal",
        axis.title = element_text(size= 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size=12),
        plot.title = element_text(hjust = 0.5, size = 14))
        

```

Let's save this 
```{r}
ggsave(plot = faceted_plot_wearlog2, filename ="thesis_facetedplot.png", dpi =600, width = 12, height = 14, path ="D:/cyepi/code/outputs")
ggsave(plot = faceted_plot_wearlog2, filename ="thesis_facetedplot.pdf", dpi =600, width = 12, height = 14, path ="D:/cyepi/code/outputs")

```

##Now let's look at individual participants
We identified 201 as the participant with most wearlog entries and participant 213 with the least. 
```{r}
mostentries_leastentries <- stacked_overview_long %>%
  filter(record_id == "201" | record_id == "214")


mostentries_leastentries_p <- ggplot(mostentries_leastentries, aes(x = count, y = factor(event, custom_order_stacked), fill = time)) +
  geom_bar(position = "stack", stat = "identity")  +
  facet_wrap(~record_id, nrow = 1, scales = "free_x") + 
  scale_y_discrete(breaks = custom_order_stacked, labels = c("Re-entering Tübingen", "Leaving Tübingen", "Off and black bag not used", "Off and black bag used", "Off before sleep ", "Off", "On", "Button press", "All wear log entries")) +
scale_fill_manual(name = "Time of reporting", values = c("#66CCFF", "#0066CC","#006633", "#CC66FF", "#FF99FF", "#FF9933", "#CC0033", "#009933"), labels = c( "Current", "Retrospective", "z1", "z2", "z3","z4", "z5", "z6" )) +
   ggtitle(bquote("Comparison between n=2 participants")) +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold")) +
  xlab("Count") +
  ylab("Event") +
   theme_bw() +
        theme() +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = "top",
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, size = 14, face = "plain"),
        axis.title = element_text(size= 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size=12),
        axis.text = element_text(size=12))

        
```

```{r}
ggsave(plot = mostentries_leastentries_p, filename ="most_leastentries.png", dpi =600, width = 12, height = 7, path ="D:/cyepi/code/outputs")
ggsave(plot = mostentries_leastentries_p, filename ="most_leastentries.pdf", dpi =600, width = 12, height = 10, path ="D:/cyepi/code/outputs")
```

```{r}
library(cowplot)

multiplot <- plot_grid(summary_plot_all, mostentries_leastentries_p, labels=c("A","B"), nrow=2, rel_heights = c(3,3))

ggsave(plot = multiplot, filename ="wearlogmultiplot.png", dpi =600,  width = 9, height = 9, path ="D:/cyepi/code/outputs")
ggsave(plot = multiplot, filename ="wearlogmultiplot.pdf", dpi =600, width = 9, height = 9, path ="D:/cyepi/code/outputs")
```

