---
title: "nonwear_filtering"
author: "Carolina Guidolin"
date: "2023-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Taking out non-wear times fro the light logging dataset 
We have identified periods of non-wear time according to the wear log. Now, we'd like to substitute these periods with NA values for each participant
```{r}
library(tidyverse)
library(hms)
library(here)
library(scales)
library(lubridate)
library(LightLogR)
```


#### Loading actlumus data 
First, we need to import the ActLumus file where the button press information is contained. We will need this later.
```{r}
path <- "D:/cyepi/code/pre-processing/nonwear_detection/actlumus"
tz <- "Europe/Berlin"

# Get a list of file paths in the specified directory
actlumusfiles <- list.files(path, pattern = ".txt", full.names = TRUE)

# Create an empty dataframe to store the combined data
dataset.LL.all <- data.frame()

# Iterate over each file and import data
dataset.LL.all <- import$ActLumus(actlumusfiles, tz = tz, auto.id = "^(\\d{3})")

```

Every participant started at different times. We will have to specify start and end time manually
```{r}
filtered_time <- data.frame(
  Id = c("201", "202", "204", "205", "206", "208", "209", "210", "212", "213", "214", "215", "216", "218", "219", "221", "222", "223", "224", "225", "226", "227", "228", "229", "230", "231"),
  start = as.POSIXct(c("2023-08-14 12:40:00", "2023-08-14 13:26:00", "2023-08-14 14:43:00", "2023-08-28 15:42:00", "2023-08-28 14:16:00", "2023-09-04 12:36:00", "2023-09-04 14:01:00", "2023-09-04 15:20:00", "2023-09-11 13:03:00", "2023-09-11 15:06:00", "2023-09-25 14:56:00", "2023-10-02 11:30:00", "2023-10-02 14:31:00", "2023-10-16 11:18:00", "2023-10-16 12:40:00", "2023-10-23 11:22:00", "2023-10-23 13:10:00", "2023-10-30 11:34:00", "2023-10-30 13:24:00", "2023-10-30 15:05:00", "2023-10-30 16:34:00", "2023-10-30 17:44:00", "2023-11-06 11:02", "2023-11-06 12:34", "2023-11-06 14:02", "2023-11-06 17:12" )),    # Add start times
  end = as.POSIXct(c("2023-08-20 23:59:59", "2023-08-20 23:59:59", "2023-08-20 23:59:59", "2023-09-03 23:59:59", "2023-09-03 23:59:59", "2023-09-10 23:59:59", "2023-09-10 23:59:59", "2023-09-10 23:59:59", "2023-09-17 23:59:59", "2023-09-17 23:59:59", "2023-10-01 23:59:59", "2023-10-08 23:59:59", "2023-10-08 23:59:59", "2023-10-22 23:59:59" , "2023-10-22 23:59:59", "2023-10-29 23:59:59", "2023-10-29 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-05 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59", "2023-11-12 23:59:59"))            # Add end times
)

dataset.LL.all <- dataset.LL.all %>%
  inner_join(filtered_time, by = "Id") %>%
  filter(Datetime >= start, Datetime <= end) %>%
  select(-start, -end)

```


### Using the LightLogR approach, let's calculate non-wear times

First, we need to importa the data from the wear log
```{r}
filepath <- here("D:/cyepi/code/pre-processing/nonwear_detection/wearlog")

# Get the files names from directory
wearlogfiles = list.files(filepath, pattern="*.csv", full.names = TRUE)

```

Now, we calculate the non-wear times. For more details, see the file nonwear_wearlog
```{r pressure, echo=FALSE}
offtime_intervals <- 
  #filenames:
  wearlogfiles %>% 
  #import.Statechanges creates already much of the necessary structure
  LightLogR::import.Statechanges(
    sep = ";", dec = ",", Datetime.format = "dmyHM", tz = "Europe/Berlin", 
    ID.colname = record_id,
    State.colnames = 
      c("wearlog_on", "wearlog_off", "wearlog_past_on", "wearlog_past_off", "wearlog_bed", "wearlog_past_sleep"),
    State.encoding = 
      c("1", "0", "1", "0", "2", "2")
    ) %>% 
  #recoding and filtering has to be done manually
  mutate(State = case_when(
      (State == 1 | State == 2) & lag(State) == 0 ~ "on",
      State == 0 & lag(State) == 1 ~ "off",
      .default = NA_character_
  )) %>% 
  filter(!is.na(State)) %>%
  #creating the intervals
  LightLogR::sc2interval(full = FALSE, length.restriction = 7 * (60*60*24)) %>%
  filter(!is.na(Interval)) 

#Filter the original dataset and replace nonwear time with NA values
data.LL.filtered <- dataset.LL.all %>%
  interval2state(offtime_intervals) %>%
  mutate(MEDI = case_match(State, "off" ~ NA, 
                           .default = MEDI)) %>%
  select(-State)

##Lastly, let's aggregate the data in 5 minutes bins

data.LL.filtered <- data.LL.filtered %>%
  aggregate_Datetime(Datetime.colname = "Datetime",
                     unit = "5 minutes",
                     numeric.handler = mean)

```

###Old code
```{r}
#Filter the original dataset and replace nonwear time with NA values
dataset.LL.filtered <- dataset.LL.all %>%
  group_by(Id) %>%
  full_join(offtime_intervals) %>%
  mutate(ind = (Datetime >= off_start & Datetime <= off_end),
         MEDI = ifelse(ind, NA, MEDI)) %>%
  distinct(Id, Datetime, .keep_all = TRUE) %>%
  ungroup() %>%
  select(-ind)

#Check that it looks as it is supposed to 
data.LL.filtered_check <- data.LL.filtered %>%
  filter(is.na(MEDI))

###Old code below:
  #filter out intervals we do not like
  filter(State == "off") 

  #calculate the nonwear time through the interval column
  mutate(off_start = int_start(Interval)) %>% 
  mutate(off_end = int_end(Interval)) %>%
  select(-Interval)
  
```

