---
title: "non-wear time preprocessing"
author: "Carolina Guidolin"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data preprocessing step 1b: detection of non-wear time

## Calculation of wear and non- wear time according to the Wear log

Here, we'd like to provide a duration of non-wear time according to the information the participants enter in their "Wear log" questionnaire every day. 

#### Loading packages 
```{r}
library(tidyverse)
library(ggplot2)
library(purrr)
library(tidyr)
library(hms)

```

#### Loading in the wearlog data for all participants and merge it into a single dataframe
All participants until 2023.11.10, meaning n = 22
```{r}

files <- setwd("D:/cyepi/code/pre-processing/nonwear_detection/wearlog")

# Get the files names from directory
wearlogfiles = list.files(pattern="*.csv")

# First apply read.csv, then rbind
wearlog_df = do.call(rbind, lapply(wearlogfiles, function(x) read.csv(x, stringsAsFactors = FALSE, sep = ";")))
```
### Calculate nonwear time
First, we want to convert the wearlog information in long format. We also want to select the relevant columns for calculation of nonwear time. For example, here we don't care whether a participant left Tübingen or not, whether they put the glasses in the bag etc. We just want to keep the info from when they took the light glasses on, when they put them on, and when they took them off before sleep. We get this info both from "current" events and "past events". 

```{r}
wearlog_df_long <- wearlog_df %>%
  select("record_id", "wearlog_on", "wearlog_off", "wearlog_past_on", "wearlog_past_off", "wearlog_bed", "wearlog_past_sleep") %>%
  group_by(record_id) %>%
  pivot_longer(cols = c("wearlog_on", "wearlog_off", "wearlog_past_on", "wearlog_past_off",  "wearlog_bed", "wearlog_past_sleep"),
               names_to = "event",
               values_to = "timestamp")

```

Second, we want to assign NA values to the blank cells and eliminate them. These cells are created when transforming the data from wide to long format, so we do not need them. 
```{r}
#Create function to to this (took this from Stack overflow)
empty_as_na <- function(x){
    if("factor" %in% class(x)) x <- as.character(x) ## since ifelse wont work with factors
    ifelse(as.character(x)!="", x, NA)
}

#Apply function to the dataset
wearlog_df_long <- wearlog_df_long %>%
  mutate_each(funs(empty_as_na))

#Filter out the NAs
wearlog_df_long <- wearlog_df_long %>%
  filter(! is.na(timestamp))
```

Third, we want to create a ternary signal where the "on" events are coded as 1, the "off" events are coded as 0 and "off before sleep" is coded as 2. 
```{r}
wearlog_df_long <- wearlog_df_long %>%
  mutate(event_ternary = case_when(
    event %in% c("wearlog_on", "wearlog_past_on") ~ 1,
    event %in% c("wearlog_off", "wearlog_past_off") ~ 0,
    event %in% c("wearlog_bed", "wearlog_past_sleep") ~ 2,
    TRUE ~ NA  # Handle other cases as needed
  ))
```

Fourth, we now want to turn the timestamp column into a POSIXct class
```{r}
#Currently, the date timestamp column has format dd.mm.yyyy with dot rather than -, so first we change that
wearlog_df_long <- wearlog_df_long %>%
  mutate(timestamp = gsub("\\.", "-", timestamp))

#Here, we clean the timestamp column using trimws() and then turn the timestamp clumn in a column called datetime, which is POSIXct 
wearlog_df_long <- wearlog_df_long %>%
  mutate(timestamp = trimws(timestamp),
         datetime = dmy_hm(timestamp))

```

Fifth, we need to sort the datetime column chronologically for each record_id
```{r}
#We remove the timestamp colum (character)
wearlog_df_long <- wearlog_df_long %>%
  select(-timestamp) 

#We use arrange() to sort the timestamps chronologically
wearlog_df_long <- wearlog_df_long %>%
  group_by(record_id) %>%
  arrange(datetime, .by_group = TRUE) 
```

Sixth, we'd like to insert some quality checks for our dataset now: 
A) The first event for every participant has to be an "on" (1) event. 
```{r}
wearlog_qualitycheck <- wearlog_df_long %>%
  group_by(record_id) %>%
  group_walk(~{
    if (first(.x$event_ternary) != 1) {
      stop(paste("Error: first event not ON for record_id", first(.x$record_id)))
    }
  })
## I think this is a rather complex way of doing what I am trying to do. Maybe a for loop would work better?
```

B) There are no two consecutive events, e.g. no 2 "on" events logged one after the other (this would not make sense)
```{r}
error_messages <- character(0) # We create an empty vector where we can store the errors

# We iterate through the data frame starting from the second row until the second to last row
for (i in 2:(nrow(wearlog_df_long) - 1)) {
  if (wearlog_df_long$event_ternary[i] == wearlog_df_long$event_ternary[i - 1]) { # we check if the current row has the same value as the previous row
    # If that's the case, add an error message
    error_messages <- c(
      error_messages,
      paste("Consecutive values at record_id:", wearlog_df_long$record_id[i],
            "datetime:", wearlog_df_long$datetime[i])
    )
  }
}

# Check if there are error messages
if (length(error_messages) > 0) {
  # There are errors; print the error messages
  for (error_message in error_messages) {
    cat("Error:", error_message, "\n")
  }
} else {
  cat("No consecutive value errors found.")
}
```
We got around 10 errors. Going back to the original file (csv downloaded from REDCap), I check was was going on at these timestamps. Here a brief description and how I adjusted it manually:

1. Participants accidentally reported same event one after the other. In this case, the first entry was kept.
2. Participants forgot to report the first ON event in the morning. In this case, the timestamp was given sing the sleep diary information (at what time the participant woke up). Although this is not 100% accurate, participants are instructed to wear the light glasses after wake up, so it is the best approximation.
3. Participants did not correct for the date when logging an "off before sleep" event which was past midnight. This means that when arranging the timestamps chronologically, this event got placed elsewhere than intended.

After correcting for these and running the code again, we still encounter 1 error for non-alternating values at the first entry for participants 226.Checking the df, it is clear that this comes from the respective previous participant 225 having two "on" values as their last value. We remove these with the code below:

```{r}
# Delete last 2 rows of participant 226
wearlog_df_long <- wearlog_df_long %>%
  filter(!(record_id == 226 & row_number() > (n() - 2)))
```


Finally, we want to calculate the duration between on an ON and an OFF event.To do this, we make two dataframes: 

1) on_events contains all the "on" events. These are defined by event_ternary == 1 (on) OR == 2 and the previous entry being a "off" entry (lag(event_ternary) == 0). The reason why we treat event_ternary ==2 ("off before sleep") as an "on" entry is because if participants remove the glasses at 22:20 and then place them on the bedside table before sleep at 22:50, that time is considered non-wear time. Similarly, we need to make sure that we count "on" events any entry which is ONLY preceded by an "off" entry. For example, if an "on" entry is preceeded by an "off before sleep" entry, I don't want to be calculating this as non-wear time, since it's actually just "sleep time".

2) off_events contains all the "off" events. These are defined as "off" entries (event_ternary == 0) AND the previous entry being an "on" event ((lag(event_ternary) == 1)).

```{r}

on_events <- wearlog_df_long %>%
  filter((event_ternary == 1 | event_ternary == 2) & lag(event_ternary) == 0) %>%
  select(record_id, datetime) %>%
  rename("off_end" = datetime)

off_events <- wearlog_df_long %>%
  filter(event_ternary == 0 & (lag(event_ternary) == 1)) %>%
  select(record_id, datetime) %>%
  rename("off_start" = datetime)

```

OK, we now want to bring these 2 dataframes together
```{r}
nonwear_events <- bind_cols(on_events, off_events) %>%
  select(-record_id...3) %>% #get rid of this self-originated column 
  rename("record_id" = record_id...1) %>% #rename this column to the original name
  mutate(nonwear_time = difftime(off_end, off_start, units = "mins", tz = tz)) %>% #calculate the difference between end of an off event and start of the off event
  mutate(nonwear_time = as_hms(nonwear_time)) %>% #turn this into a POSIXct object
  mutate(date = as.Date(off_end)) # since ultimately care about the daily nonwear, we want to extract the date from the POSIXct 

```
Now that we have calculated all the non-wear times, we want to sum them up for every day. In other words, we want to calculate the daily non-wear of each participant. 
```{r}
nonwear_df_daily <- nonwear_events %>%
  group_by(record_id, date) %>%
  mutate(nonwear_daily = sum(nonwear_time)) %>%
  mutate(nonwear_daily = as_hms(nonwear_daily)) %>% # turn it into a POSIXct object
  select(-c(off_end, off_start, nonwear_time)) %>%
  ungroup()

```
Note that the caveat of this method is that it doesn't calculate the total non-wear time according to the individual's day (sleep/wake time), but rather according to the 24-h day starting at 00:0 and ending at 00:00 of the next day. 

This works, but we still have multipe entries for every date. Instead, we just need one value representing the daily non-wear for each day for each participant. Also, since we don't necessarily care about the dates but rather days of the week, we can turn dates into days 
```{r}
nonwear_df_daily <- nonwear_df_daily %>%
  group_by(record_id) %>%
  distinct(date, .keep_all = TRUE) %>%
  ungroup() 

Sys.setlocale("LC_TIME", "en_US.UTF-8") #needed to run this code to have the days displayed in English and not German
nonwear_df_daily$day <- format(nonwear_df_daily$date, format = "%A", locale="English")
```

And now, it's time to plot the non-wear time according to the wear log!
```{r}
nonwear_df_daily$day <- factor(nonwear_df_daily$day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) #use the factor function with levels specified so that the plot will show the days of the week in order, i.e. Monday to Sunday

ggplot(nonwear_df_daily, aes(x = day, y = nonwear_daily)) +
  geom_violin(alpha = 0.3, aes(fill=day), trim = FALSE) + 
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold")) +
  xlab("Experimental day") + 
  ylab("Non-wear time (hours)") +
  ggtitle("Participants' non-wear time during the experimental week\n according to self-reported wear log entries") 
  
```

It looks like the range of weekly non-wear time for most participants is something between 00:00:00 and 04:00:00. At the weekend, it looks like this increases slightly for some participants. We also have some outliers, where non-wear time seems to be very high (>4hrs). Let's filter the dataset to check who these participants are.
```{r}
outliers <- nonwear_df_daily %>%
  filter(nonwear_daily >= as.difftime(4, units = "hours")) %>% #since nonwear_daily is of difftime hms class, we need to use as.difftime to compare this to 
  select(record_id, day, nonwear_daily)
```

After visual inspection, I can say that for participants 201, 204, 208, 215, 221 and 223, the wear log data is correct and they just did not wear the light glasses for a long time. For example, they took them off in the evening at around 22:00, then went partying (information from other questionnaires where participants are asked about their activity) and came back late and placed the light glasses on the nightstand in the early hours of the morning. However, for participant 224 (non-wear time on Sat ~14hrs) probably something went wrong with the logging of wear time on the wear log (e.g. participant forgot to report putting the glasses back on). For now we will leave it as such and hope to fix this when looking at other sources of non-wear.




  
