---
title: "bag_activity_prc"
author: "Carolina Guidolin"
date: "2024-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Note
This script should be run after files in 01_import and 02_datapreparation have been run. 

# Aim
To know whether we can precisely predict non-wear times (detected by the wear log) using information from the black back use (low mEDI) and/or activity (low PIM). To do this, we build a precision recall curve (PRC).

## Algorithm which indentifies clusters of non-wear time 
These functions (find_clusters_funs, loaded in separately) identify clusters of low illuminance/activity (or an arbitrary variable). The following parameters can be selected: 
1) logical values where clusters should be found - in this case, low illuminance or low activity values (below a certain threshold)
2) minimum length of the clusters
3) max interrpution length for the cluster to still be identified as such 

## Before running the cluster detection algorithm
First, let's explore the relationship between light and activity across wear log status. This visualisation step helps to understand whether it makes sense to even detect low activty or illuminance as a proxy of non-wear. 
```{r}
#We will be plotting using log10 transformation, so we cannot have 0 values in our data set. For this reason, we add 1 to each value in MEDI and PIM

# Create a dataset for plotting
vis.LL.wrlg <- dataset.LL.wrlg

# Add a value of 1 to the relevant columns
vis.LL.wrlg$MEDI <- vis.LL.wrlg$MEDI + 1
vis.LL.wrlg$PIM <- vis.LL.wrlg$PIM + 1

#To be able to visually observe how light and activity relate to each other in our data set, we cannot plot all our data points (too many). So, we take median values of each Id, for each State (off/on/sleep)

plot.median.LL.wrlg <- vis.LL.wrlg %>%
#  filter(!State == "sleep") %>%
  group_by(Id, State) %>%
  summarise(median_medi = median(log10(MEDI), na.rm = TRUE),
            median_pim = median(log10(PIM), na.rm = TRUE),
            qr1_medi = quantile(log10(MEDI), 1/4, na.rm = TRUE), #first quantile
            qr1_pim = quantile(log10(PIM), 1/4, na.rm = TRUE), #first quantile
            qr3_medi = quantile(log10(MEDI), 3/4, na.rm = TRUE), #third quantile
            qr3_pim = quantile(log10(PIM), 3/4, na.rm = TRUE)) %>% #third quantile
  ungroup()

# Plot
set.seed(1) #we will use geom_jitter, which uses random sampling. For reproducibility we need to set.seed(1) so that running the script multiple times will lead to the same plot

scatter <- ggplot(plot.median.LL.wrlg, aes(x = median_medi, y = median_pim)) + 
  geom_jitter(aes(fill = State, color = State), size = 2.5, alpha = 0.6,
              position = position_jitter(width = .3, height = .3)) + #adjust position of points to avoid overlap
   scale_fill_manual(
    name = "Wear log \nstate",
    values = c("off" = "darkred", "on" = "#0072B2", "sleep" = "darkgoldenrod2"),
    labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  scale_color_manual(
    name = "Wear log \nstate",
    values = c("off" = "darkred", "on" = "#0072B2", "sleep" = "darkgoldenrod2"),
    labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  scale_x_continuous(
    breaks = c(-1, 1,3), 
    labels = c(expression(10^-1), expression(10^1), expression(10^3)), #ensure that the log expression is labelled correctly 
    limits = c(-1, 4)
  ) +
  scale_y_continuous(
    breaks = c(-1, 1, 3),
    labels = c(expression(10^-1), expression(10^1), expression(10^3)), #ensure that the log expression is labelled correctly 
    limits = c(-1, 4)
  ) +
  labs(x = "Median of log illuminance (mEDI, lx)",
       y = "Median of log activity (PIM)") +
  coord_fixed(ratio=1) + #make sure plot is a square
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12, hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1)) 
  

marginal <- ggExtra::ggMarginal(scatter, type = "boxplot", groupColour = TRUE, groupFill = TRUE)


plot_marginal <- ggpubr::annotate_figure(marginal,
                                 #Add title
                                 top = text_grob("Illuminance and activity \nacross wear log states",
                                                            color = "black",
                                                            face = "plain",
                                                            size = 18)) 

plot_marginal <- plot_marginal + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

plot_marginal

```
### Saving the plot
```{r}

ggsave(filename = "results_fig07a.png",
       plot = plot_marginal ,
       width = 5,
       height = 5.5,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "fig07"))
```

### Conclusion from this plot
It seems like during wear time, activity and light are high, while during non-wear time, they are low. Hence, we will next implement a method to detect clusters of low illuminance and low activity as a way of classifing non-wear time. 

### Calculating descrptive metrics from this plot
We want to know range of median values for each Wear log state
```{r}
#Because we had added a 1 to the original dataset for plotting, we want to revert this back when calculating metrics - so we recalculate median.LL.wrlg directly on dataset.LL.wrlg
calc.median.LL.wrlg <- dataset.LL.wrlg %>%
  group_by(Id, State) %>%
  summarise(median_medi = median(MEDI, na.rm = TRUE),
            median_pim = median(PIM, na.rm = TRUE)) %>%
  ungroup()

#Now, we can calculate some descriptive statistics
medi_pim_bystate <- calc.median.LL.wrlg %>%
  group_by(State) %>%
  summarise(mean_medi = mean(median_medi),
            sd_medi = sd(median_medi),
            mean_pim = mean(median_pim),
            sd_pim = sd(median_pim),
            max_medi = max(median_medi),
            min_medi = min(median_medi),
            max_pim = max(median_pim),
            min_pim = min(median_pim),
            median_medi = median(median_medi),
            median_pim = median(median_pim),
            range_medi = max_medi-min_medi,
            range_pim = max_pim-min_pim) %>%
  ungroup()

```


### Creating a dataframe without sleep states
Since we want to classify wear and non-wear based on clusters of low illuminance, we do not care about the time that the participants are sleeping. In other words, the wear and non-wear classification does not apply to this period. To be able to use the dataset.LL.wrlg and run the classification on it, we first need to set all "sleep" values (based on Wear log data) as NA
```{r}
df.LL.nosleep <- dataset.LL.wrlg %>%
  mutate(State = if_else(State == "sleep", NA_character_, State)) 
```

## Proceed to apply cluster detection algoirthm for low activity (PIM)

### First, import functions
```{r}
# Fuction to identify clusters
base::source(here::here("03_analysis", "functions", "find_clusters_fun.R"))

#Function to generate aggregate pr curve
base::source(here::here("03_analysis", "functions", "generate_prc_funs.R"))

#Function to generate individual pr curve
base::source(here::here("03_analysis", "functions", "generate_indiv_prc_funs.R"))
```


### Identifying clusters of low activity (PIM)
```{r}

#Create a sequence of PIM thresholds at 5 unit steps
pimthresholds <- c(5,10, 15, 20, 25, 30, 35, 40, 45, 50)

#Empty list to store classification results
prc_list <- list()

#Running a for loop to generate a prc for all selected PIM thresholds 
for (threshold in pimthresholds) {
  
  prc_result <- generate_prc(
    dataset = df.LL.nosleep,
    low_var = "PIM",
    min_length = 72, #12 minutes (n of observations)
    max_interrupt = 0, #0 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_list[[as.character(threshold)]] <- prc_result
  
}

#Turn into a df
prcurve_pim <- bind_rows(prc_list)

#Turn threshold to numeric for plotting
prcurve_pim$threshold <- as.numeric(prcurve_pim$threshold)

```

### Calculating F1 score
This scores is a metric to evaluate precision and recall. It is calculate as the harmonic mean of precision and recall. The F1 Score is useful when looking for the right balance between high precision and high recall, as it penalizes extreme negative values of either component. 
```{r}

prcurve_pim_f1 <- prcurve_pim %>%
  mutate(f1_score = (2 * PPV * TPR)/(PPV + TPR))

```

## Proceed to apply cluster detection algorithm to low light levels (mEDI)
### Identifying clusters of low light levels (mEDI)
```{r}

medithresholds <- seq(1, 10, by = 1)

#Empty list to store classification results
prc_list <- list()

for (threshold in medithresholds) {
  
  prc_result <- generate_prc(
    dataset = df.LL.nosleep,
    low_var = "MEDI",
    min_length = 126, #21 minutes (n of observations)
    max_interrupt = 0, #0 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_m_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_list[[as.character(threshold)]] <- prc_result
  
}

#Turn list into df
prcurve_medi <- bind_rows(prc_list)
#Turn threshold to numeric for plotting
prcurve_medi$threshold <- as.numeric(prcurve_medi$threshold)


```

### Calculate F1 score 
```{r}

prcurve_medi_f1 <- prcurve_medi %>%
  mutate(f1_score = (2* PPV * TPR)/(PPV + TPR))

```


### Visualisation of the precision recall curves for activity and light (all participants together)
```{r}

prc_combined <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  #First plot: activity
  geom_point(data = prcurve_pim, aes(x = TPR, y = PPV, color = threshold)) +
   scale_color_gradient(
     name = "Activity threshold (PIM)",
     low = "blue", high = "lightblue",
     limits= c(5,50),
    guide = guide_colorbar(title.position = "top",
                           title.hjust = 0.5))  +
  ggnewscale::new_scale_color() + #need this to reset the color scale for the next plot
  #Adding the arrow to show directionality (for illuminance)
  geom_segment(aes(x = 0.84,
                   y = 0.78,
                   xend = 0.92,
                   yend = 0.38),
               arrow = arrow(length=unit(.2, "cm")), lwd = 0.8) +
  #Second plot: light
  geom_point(data = prcurve_medi, aes(x = TPR, y = PPV, colour = threshold)) +
   scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(1,10),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  #Adding the arrow to show directionality (for activity)
  geom_segment(aes(x = 0.59,
                   y = 0.42,
                   xend = 0.74,
                 yend = 0.33),
               arrow = arrow(length=unit(.2, "cm")), lwd = 0.8) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)") +
  #Adding the symbols for the arrows as text and the label
  annotate("text",
           x = 0.765,
           y = 0.84,
           label = sprintf('\u2191')) + #arrow pointing upwards for Illuminance
  geom_text(aes(x= 0.9,
                y = 0.83,
                label = "Illuminance")) + #label of illuminance
  annotate("text",
           x = 0.40,
           y = 0.46,
           label = sprintf('\u2191')) + #arrow pointing upwards for activity
  geom_text(aes(x = 0.49,
                y = 0.45,
                label = "Activity")) + #label for activity
  geom_text(aes(x=0.215,
                y = 0.535,
                label = "Baseline classifier"),
            color = "darkgrey",
            size = 3.8) +
  theme_bw() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 8),
        legend.key.size = unit(4, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal",
        plot.margin = margin(t=0,
                                   b=0,
                                   r=0,
                                   l=0)) +
  coord_fixed(ratio = 1) 



#Save the plot
ggsave(filename = "results_fig07b.png",
       plot = prc_combined,
       height = 5.5, 
       width = 5,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "fig07"))
```

### Combining the plots
```{r}
library(cowplot)

prc_combined_titled <- ggpubr::annotate_figure(prc_combined,
                                 #Add title
                                 top = text_grob("Precision recall curve for \ndetection of non-wear intervals",
                                                            color = "black",
                                                            face = "plain",
                                                            size = 18,
                                                 just = 0.35)) 

prc_combined_titled <- prc_combined_titled + theme(plot.margin = unit(c(0, 0, 0, 0), "cm"))

medi_pim_fig <- cowplot::plot_grid(plot_marginal,
                          prc_combined_titled,
                          labels = c("A", "B"),
                          nrow = 1,
                          align = "hv",
                          axis = "tblr")
# Save the plot
ggsave(filename = "results_fig07_all.png",
       plot = medi_pim_fig,
       height = 5.5, 
       width = 8.9,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "fig07"))

```

##Visualisation of activity and light prc for each participant in a multiplot
This is helpful to know whether the prc performs differently in each participant. We also want to know: if our algorithm performs poorly, is it because the participants were not compliant with using the black bag? So, the first step is actually to calculate how often the bag was actually used when logging and off event. 

###Calculating % usage of black bag 
```{r}
 bag_use <- df.LL.nosleep %>%
  group_by(Id) %>%
   mutate(bag_use = case_when(
     State == "off" & bag == 0 ~ "notused", #when State is off and bag is 0, it was not used
     State == "off" & bag == 1 ~ "used", #when state is off and bag is 1, it was used
     .default = NA_character_)) %>%
  ungroup()

per_bag_use <- bag_use %>%
  group_by(Id, bag_use) %>% 
  summarise(count = n()) %>% #calculating how many used and notused per participant
  filter(!is.na(bag_use)) %>% #eliminate NAs which correspond to when State is on or sleep
  complete(bag_use = c("used", "notused"), fill = list(count = 0)) %>% #if participant does not have values for "used" or "notused" because they either never used the bag, or always used it, we replace that with 0
  mutate(total = sum(count), #counting total entries per participant
         compliance_per = if_else(bag_use == "used", (count*100)/total, NA), #calculating % of "used" over total entries
         compliance_per = round(compliance_per, 1)) %>% #rounding the percentage value to 1 decimal point
  filter(!is.na(compliance_per)) #eliminate the NA that comes from the ifelse

```

## Plotting the PRC for each participant and labelling % compliance of bag use 
For this, we need to run the prc curves with the correct function (generate_indiv_prc)

## Run this for activity first
```{r}
# Generating prc for each individual using activity 

#Create a sequence of PIM thresholds at 5 unit steps
pimthresholds <- c(5,10, 15, 20, 25, 30, 35, 40, 45, 50)

#Empty list to store classification results
prc_list <- list()

#Running a for loop to generate a prc for all selected PIM thresholds 
for (threshold in pimthresholds) {
  
  prc_result <- generate_indiv_prc(
    dataset = df.LL.nosleep,
    low_var = "PIM",
    min_length = 72, #12 minutes (n of observations)
    max_interrupt = 0, #0 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_list[[as.character(threshold)]] <- prc_result
  
}

#Turn into a df
prcurve_pim_indiv <- bind_rows(prc_list)

#Turn threshold to numeric for plotting
prcurve_pim_indiv$threshold <- as.numeric(prcurve_pim_indiv$threshold)
```

## Run this for illuminance

```{r}

medithresholds <- seq(1, 10, by = 1)

#Empty list to store classification results
prc_list <- list()

for (threshold in medithresholds) {
  
  prc_result <- generate_indiv_prc(
    dataset = df.LL.nosleep,
    low_var = "MEDI",
    min_length = 126, #21 minutes (n of observations)
    max_interrupt = 0, #0 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_m_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_list[[as.character(threshold)]] <- prc_result
  
}

#Turn list into df
prcurve_medi_indiv <- bind_rows(prc_list)
#Turn threshold to numeric for plotting
prcurve_medi_indiv$threshold <- as.numeric(prcurve_medi_indiv$threshold)
```

## Plotting each participant in a panel
```{r}
prc_multiplot <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  #First plot: activity
  geom_point(data = prcurve_pim_indiv, aes(x = TPR, y = PPV, color = threshold)) +
   scale_color_gradient(
     name = "Activity threshold (PIM)",
     low = "blue", high = "lightblue",
     limits= c(5,50),
    guide = guide_colorbar(title.position = "top",
                           title.hjust = 0.5)) +
  scale_x_continuous(breaks = c(0.0, 0.50, 1.0),
                     labels = c("0", "0.50", "1"),
                     limits = c(0, 1)) +
      coord_fixed(ratio = 1) +
  facet_wrap(~ Id, scales = "fixed") +
  ggnewscale::new_scale_color() +
  #Second plot: light
  geom_point(data = prcurve_medi_indiv, aes(x = TPR, y = PPV, colour = threshold)) +
   scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(1,10),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~Id, , scales = "fixed") +
  geom_text(data = per_bag_use, aes(x = 0.15, y = 0.90, label = paste(compliance_per,"%")), size = 3.3) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value (precision)",
       title = "Precision-recall curve for detection of non-wear intervals",
       ) +
   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 16, colour = "grey33"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        axis.title.x = element_text(hjust = 0),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 10),
        legend.key.size = unit(6, "mm"),
        legend.position = "top",
        legend.box = "horizontal") +
   coord_fixed(ratio = 1)

#Save the plot
ggsave(filename = "results_fig08.png",
       plot = prc_multiplot,
       height = 11, 
       width = 10,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "fig08"))

```

