---
title: "metrics_comparison"
author: "Carolina Guidolin"
date: "2024-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Calculating metrics using different datasets
This script should be run after classification_summary. 

The aim of this script is to calculate and compare light metrics on three different datasets:
1. A raw dataset, where non-wear information is ignored (i.e., we include all data regardless of whether it is wear or non-wear)
2. A dataset where non-wear information is identified using self-reported non-wear time (wear log)
3. A dataset where non-wear information is identified by detecting clusters of low illuminance

### Creation of the three datasets
Before performing the comparison, we need to identify these three datasets. 
```{r}
# 1. Raw dataset: this is the raw data, i.e. dataset.LL.wrlg
raw_dataset <- dataset.LL.wrlg

# 2. Non-wear by wear log dataset. We need to create this from dataset.LL.wrlg
nw_wrlg <- dataset.LL.wrlg %>% 
  mutate(MEDI = if_else(State == "off", NA, MEDI))

#3. Non-wear by detection of clusters of low illuminance. We will use the dataset where padding has been added to transition states (see script classification_summary.Rmd). 
nw_clusters <- nw_alg_padded %>%
  mutate(MEDI = case_when(
    is.na(is_low_medi_cluster) ~ MEDI, # When is_low_medi_cluster is NA, retain MEDI
    is_low_medi_cluster == 0 ~ NA, # When is_low_medi_cluster is 0, set MEDI to NA
    TRUE ~ MEDI # For all other cases, retain MEDI
  ))

```

### Data prep
```{r}
# Exclude Monday since this day does not contain full data. Do this for all three data frames
raw_dataset <- raw_dataset %>% dplyr::filter(weekdays(Datetime) != "Monday")
nw_wrlg <- nw_wrlg %>% dplyr::filter(weekdays(Datetime) != "Monday")
nw_clusters <- nw_clusters %>% dplyr::filter(weekdays(Datetime) != "Monday")
```

### 1. Calculating interdaily stability 
```{r}
# 1a. raw dataset
is_raw_dataset <- raw_dataset %>% 
  summarize(
    IS_raw = interdaily_stability(
      Light.vector = MEDI,
      Datetime.vector = Datetime
    )
  )

# 1b. With Wear log based non-wear 
is_nw_wrlg <- nw_wrlg %>% 
  summarize(
    IS_wrlg = interdaily_stability(
      Light.vector = MEDI,
      Datetime.vector = Datetime,
      na.rm = TRUE
    )
  )

# 1c. With cluster based non-wear
is_nw_clusters <- nw_clusters %>%
  group_by(Id) %>% #we need to group here, because this df is not directly generate in LightLogR
  # and hence the data is not grouped by Id by default (as is the case for the other 2 data frames)
  summarize(
    IS_clusters = interdaily_stability(
      Light.vector = MEDI,
      Datetime.vector = Datetime,
      na.rm = TRUE
    )
  ) %>% 
  ungroup()

# Combining the dfs 
temp <- full_join(is_raw_dataset, is_nw_wrlg, by = "Id") #first, combine raw and wrlg data sets
is_all <- full_join(temp, is_nw_clusters, by = "Id") #second, add clusters data set

# Calculating deltas
is_all <- is_all %>%
  mutate(delta_wrlg = IS_wrlg - IS_raw,
         delta_clusters = IS_clusters - IS_raw)


```

#### Plotting interdaily stability
```{r}
library(patchwork)

# Raw vs clean (wrlg) scatterplot
clean_wrlg <- ggpubr::ggscatter(data = is_all,
          x = "IS_raw",
          y = "IS_wrlg",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
  geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(0, 1),
    breaks = c(0, 0.50, 1), 
    labels = c("0", "0.50", "1")
  ) +
  scale_y_continuous(limits = c(0, 1),
    breaks = c(0, 0.50, 1), 
    labels = c("0", "0.50", "1")
  ) + 
  labs(x = "Raw dataset", y = "Clean dataset (Wear log)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Raw vs clean (clusters) scatterplot
clean_clusters <- ggpubr::ggscatter(data = is_all,
          x = "IS_raw",
          y = "IS_clusters",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
   geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(0, 1),
    breaks = c(0, 0.50, 1), 
    labels = c("0", "0.50", "1")
  ) +
  scale_y_continuous(limits = c(0, 1),
    breaks = c(0, 0.50, 1), 
    labels = c("0", "0.50", "1")
  ) + 
  labs(x = "Raw dataset", y = "Clean dataset (algorithm)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Histogram of deltas
## First, do some re-structuring
# Turning df into long form to perform computations
is_all_long <- is_all %>%
  pivot_longer(cols = c(delta_wrlg, delta_clusters),
                        names_to = "metric",
                        values_to = "delta") 


delta_hist <- ggplot(is_all_long, aes(x = delta, fill = metric)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(x = "Delta", y = "Frequency") +
  scale_fill_manual(
    name = "Metric",
    values = c("#0072B2", "#D55E00"),
    labels = c("Algorithm dataset", "Wear log dataset")
    ) +
  scale_x_continuous(breaks = c(-0.10, 0.00, 0.10),
                     labels = c("-0.10", 0, "0.10")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        aspect.ratio = 1,
        plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))  
 # ggpubr::rremove("legend")


# Combining the plots
## Make sure you have patchwork loaded in
p1 <- clean_wrlg + clean_clusters + delta_hist + 
  patchwork::plot_annotation("IS",
  tag_levels = list(c("A1", "A2", "A3")),
  theme = theme(plot.title = element_text(hjust = 0.5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p1

```

### 2. Calculating intradaily variability
```{r}
# 1a. raw dataset
iv_raw_dataset <- raw_dataset %>% 
  summarize(
    IV_raw = intradaily_variability(
      Light.vector = MEDI,
      Datetime.vector = Datetime
    )
  )

# 1b. With Wear log based non-wear 
iv_nw_wrlg <- nw_wrlg %>% 
  summarize(
    IV_wrlg = intradaily_variability(
      Light.vector = MEDI,
      Datetime.vector = Datetime,
      na.rm = TRUE
    )
  )

# 1c. With cluster based non-wear
iv_nw_clusters <- nw_clusters %>%
  group_by(Id) %>% #we need to group here, because this df is not directly generate in LightLogR
  # and hence the data is not grouped by Id by default (as is the case for the other 2 data frames)
  summarize(
    IV_clusters = intradaily_variability(
      Light.vector = MEDI,
      Datetime.vector = Datetime,
      na.rm = TRUE
    )
  ) %>% 
  ungroup()

# Combining the dfs 
temp <- full_join(iv_raw_dataset, iv_nw_wrlg, by = "Id") #first, combine raw and wrlg data sets
iv_all <- full_join(temp, iv_nw_clusters, by = "Id") #second, add clusters data set

# Calculating delta
iv_all <- iv_all %>%
    mutate(delta_wrlg = IV_wrlg - IV_raw,
          delta_clusters = IV_clusters - IV_raw)
```

#### Plotting intradaily variability 
```{r}
# Raw vs clean (wrlg)
clean_wrlg <- ggpubr::ggscatter(data = iv_all,
          x = "IV_raw",
          y = "IV_wrlg",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
   # ggpubr::stat_cor(aes(label = after_stat(r.label)), # add r coefficient but not p value
   #                method = "pearson",
    #               label.x = 0.03,
     #               label.y = 1.95) +
  geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(0, 2),
    breaks = c(0, 1, 2), 
    labels = c("0", "1", "2")
  ) +
  scale_y_continuous(
   limits = c(0, 2),
    breaks = c(0, 1, 2), 
    labels = c("0", "1", "2")
  ) + 
  labs(x = "Raw dataset", y = "Clean dataset (Wear log)") +
  coord_fixed(ratio = 1) + 
  ggpubr::rremove("legend")+
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Raw vs clean (clusters) 
clean_clusters <- ggpubr::ggscatter(data = iv_all,
          x = "IV_raw",
          y = "IV_clusters",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
    #ggpubr::stat_cor(aes(label = after_stat(r.label)), # add r coefficient but not p value
   #                method = "pearson",
    #               label.x = 0.03,
    #               label.y = 1.95) +
   geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(0, 2),
    breaks = c(0, 1, 2), 
    labels = c("0", "1", "2")
  ) +
  scale_y_continuous(
   limits = c(0, 2),
    breaks = c(0, 1, 2), 
    labels = c("0", "1", "2")
  ) + 
  labs(x = "Raw dataset", y = "Clean dataset (algorithm)") +
  coord_fixed(ratio = 1) + 
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Histogram of deltas
## First, do some re-structuring
# Turning df into long form to perform computations
iv_all_long <- iv_all %>%
  pivot_longer(cols = c(delta_wrlg, delta_clusters),
                        names_to = "metric",
                        values_to = "delta")

delta_hist <- ggplot(iv_all_long, aes(x = delta, fill = metric)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(x = "Delta", y = "Frequency") +
  scale_fill_manual(
    name = "Metric",
    values = c("#0072B2", "#D55E00"),
    labels = c("Algorithm dataset", "Wear log dataset")
    ) +
  scale_x_continuous(breaks = c(-1.5, 0, 1.5), # make sure x axis is centred around 0 
                     limits = c(-1.5, 1.5),
                     labels = c("-1.5", "0", "1.5")) + # manually selecting labels
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        aspect.ratio = 1,
        plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm")) +
  ggpubr::rremove("legend")


# Combining the plots
## Make sure you have patchwork loaded in
p2 <- clean_wrlg + clean_clusters + delta_hist + 
  patchwork::plot_annotation("IV",
  tag_levels = list(c("B1", "B2", "B3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p2

```
### Calculating batch metrics for each participants on each day of the experimental week 
- MLIT250, FLIT250, LLIT250
-  FLIT10, LLIT10
- TAT250, TAT1000
- M10m 
```{r}
raw_dataset_wdays<- raw_dataset %>%
  LightLogR::add_Time_col() %>%
  mutate(wDay = wday(Datetime, label = TRUE, week_start = 1))

nw_wrlg_wdays <- nw_wrlg %>%
  LightLogR::add_Time_col() %>%
  mutate(wDay = wday(Datetime, label = TRUE, week_start = 1))

nw_clusters_wdays <- nw_clusters %>%
  LightLogR::add_Time_col() %>%
  mutate(wDay = wday(Datetime, label = TRUE, week_start = 1))

# Batch metrics for raw dataset
batch_raw <- 
  raw_dataset_wdays %>% 
  group_by(Id, wDay) %>% 
  summarize(
    MLIT10 = #Note: This creates 1) mean timing above threshold, 2) first timing above threshold, 3) last timing above threshold
      timing_above_threshold(MEDI, Time, threshold = 10, as.df = TRUE, na.rm = TRUE),
    MLIT250 = 
      timing_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    MLIT1000 = 
      timing_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    TAT250 = 
      duration_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    TAT1000 = 
      duration_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    .groups = "drop_last"
    ) %>% 
  unnest(-Id)

# Batch metrics for nw_wrlg dataset
batch_wrlg <- 
  nw_wrlg_wdays %>% 
  group_by(Id, wDay) %>% 
  summarize(
     MLIT10 =
      timing_above_threshold(MEDI, Time, threshold = 10, as.df = TRUE, na.rm = TRUE),
    MLIT250 = 
      timing_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    MLIT1000 = 
      timing_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    TAT250 = 
      duration_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    TAT1000 = 
      duration_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    .groups = "drop_last"
    ) %>% 
  unnest(-Id)

# Batch metrics for nw_clusters dataset
batch_clusters <- 
  nw_clusters_wdays %>% 
  group_by(Id, wDay) %>% 
  summarize(
     MLIT10 = 
      timing_above_threshold(MEDI, Time, threshold = 10, as.df = TRUE, na.rm = TRUE),
    MLIT250 = 
      timing_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    MLIT1000 = 
      timing_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    TAT250 = 
      duration_above_threshold(MEDI, Time, threshold = 250, as.df = TRUE, na.rm = TRUE),
    TAT1000 = 
      duration_above_threshold(MEDI, Time, threshold = 1000, as.df = TRUE, na.rm = TRUE),
    .groups = "drop_last"
    ) %>% 
  unnest(-Id)

```

## Calculating timing metrics
In the following chunks, we will calculate "timing metrics":
- MLiT250, MLiT1000
- FLiT 10, FLiT 250, FLiT 1000
- LLit 10, LLiT 250, LLit 1000
To facilitate these calculations, we wrote a function (calculate_metric in the vis_metric_funs.R script) that takes the tree datasets and calculates mean and sd for the timing metric of interest. We also wrote a function to visualise the three plots as done for IS and IV above. Here, we first import these functions.

### Importing functions for metric calculation and viz
```{r}
# Function to calculate metrics and visualise them - same R script
base::source(here::here("03_analysis", "functions", "vis_metric_funs.R"))
```

### MLiT250 analysis
```{r}
# Calculation 
mlit250_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "mean_timing_above_250"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = mlit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = mlit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = mlit250_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p4 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("MLiT 250 lx",
  tag_levels = list(c("C1", "C2", "C3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p4
```
#### MLiT 1000 analysis
```{r}
# Calculation 
mlit1000_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "mean_timing_above_1000"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = mlit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = mlit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = mlit1000_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p5 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("MLiT 1000 lx",
  tag_levels = list(c("D1", "D2", "D3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p5
```
### FLiT 10 analysis
```{r}
# Calculation 
flit10_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "first_timing_above_10"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = flit10_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = flit10_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
   x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = flit10_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p6 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("FLitT 10 lx",
  tag_levels = list(c("E1", "E2", "E3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p6
```
### FLiT 250 analysis
```{r}
# Calculation 
flit250_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "first_timing_above_250"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = flit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = flit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = flit250_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p7 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("FLitT 250 lx",
  tag_levels = list(c("F1", "F2", "F3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p7
```

### FLiT 1000 comparison 
```{r}
# Calculation 
flit1000_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "first_timing_above_1000"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = flit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
 x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = flit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = flit1000_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p8 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("FLitT 1000 lx",
  tag_levels = list(c("G1", "G2", "G3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p8
```
### LLiT 10 analysis
```{r}
# Calculation 
llit10_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "last_timing_above_10"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = llit10_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = llit10_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = llit10_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p9 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("LLiT 10 lx",
  tag_levels = list(c("H1", "H2", "H3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p9

```
### LLiT 250 analysis
```{r}
# Calculation 
llit250_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "last_timing_above_250"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = llit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = llit250_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = llit250_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p10 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("LLiT 250 lx",
  tag_levels = list(c("I1", "I2", "I3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p10
```
### LLiT 1000 analysis
```{r}
# Calculation 
llit1000_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "last_timing_above_1000"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = llit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = llit1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = llit1000_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p11 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("LLiT 1000 lx",
  tag_levels = list(c("J1", "J2", "J3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p11
```
### TAT 250 analysis
```{r}
# Calculation 
tat250_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "duration_above_250"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = tat250_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = tat250_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = tat250_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p12 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("TAT 250 lx",
  tag_levels = list(c("K1", "K2", "K3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p12
```
### TAT 1000 comparison
```{r}
# Calculation 
tat1000_all <- calculate_metric(
  raw_df = batch_raw,
  wrlg_df = batch_wrlg,
  clusters_df = batch_clusters,
  metric = "duration_above_1000"
)

# Plotting
## Raw vs wrlg
raw_vs_wrlg_vis <- visualize_comparison(
  data = tat1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_wrlg", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (Wear log)",
)

## Raw vs clusters
raw_vs_clusters_vis <- visualize_comparison(
  data = tat1000_all, 
  x_col = "mean_raw", 
  y_col = "mean_clusters", 
  x_label = "Raw dataset", 
  y_label = "Clean dataset (algorithm)",
)

## Histogram of deltas 
delta_hist <- vis_deltas(df = tat1000_all,
           delta_col = "delta",
           metric_col = "metric") 

## Combining the plots
p13 <- raw_vs_wrlg_vis + raw_vs_clusters_vis + delta_hist +
  patchwork::plot_annotation("TAT 1000 lx",
  tag_levels = list(c("M1", "M2", "M3")),
  theme = theme(plot.title = element_text(hjust = .5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.0,b=0.1,l=0), "cm")))

p13

```
### Mean comparison 
We calculate the average daily MEDI (for this, we use the log zero inflated approach proposed [here](https://tscnlab.github.io/LightLogR/articles/log.html)). The day information is based on the given photoperiod of provided coordinates.
```{r}
#specifying coordinates (latitude/longitude)
coordinates <- c(48.521637, 9.057645)

# Raw dataset
mean_raw <- raw_dataset_wdays %>%
  add_photoperiod(coordinates) %>%
  group_by(Id, photoperiod.state, .add=TRUE) %>%
  summarise(mean_raw = exp_zero_inflated(mean(log_zero_inflated(MEDI), na.rm = TRUE))) %>%
  filter(photoperiod.state == "day") %>%
  select(-photoperiod.state)

# Wear log dataset
mean_wrlg <- nw_wrlg_wdays %>%
  add_photoperiod(coordinates) %>%
  group_by(Id, photoperiod.state, .add=TRUE) %>%
  summarise(mean_wrlg = exp_zero_inflated(mean(log_zero_inflated(MEDI), na.rm = TRUE))) %>%
  filter(photoperiod.state == "day") %>%
  select(-photoperiod.state)

# Algorithm dataset
mean_clusters <- nw_clusters_wdays %>%
  add_photoperiod(coordinates) %>%
  group_by(Id, photoperiod.state, .add=TRUE) %>%
  summarise(mean_clusters = exp_zero_inflated(mean(log_zero_inflated(MEDI), na.rm = TRUE))) %>%
  filter(photoperiod.state == "day") %>%
  select(-photoperiod.state)

# Combining the dfs 
temp <- full_join(mean_raw, mean_wrlg, by = "Id") #first, combine raw and wrlg data sets
mean_all <- full_join(temp, mean_clusters, by = "Id") #second, add clusters data set

# Calculating the delta
mean_all <- mean_all %>%
  mutate(delta_wrlg = mean_wrlg - mean_raw,
         delta_clusters = mean_clusters - mean_raw)

```

### Visualising mean 
```{r}
# Raw vs clean (wrlg) scatterplot

#Finding min max and mid values for plotting
mean_all_min_raw <- min(mean_all$mean_raw)
mean_all_min_wrlg <- min(mean_all$mean_wrlg)
mean_all_min_clusters <- min(mean_all$mean_clusters)

mean_all_max_raw <- max(mean_all$mean_raw)
mean_all_max_wrlg <- max(mean_all$mean_wrlg)
mean_all_max_clusters <- max(mean_all$mean_clusters)

mean_all_mid_raw <- (mean_all_min_raw+mean_all_max_raw)/2
mean_all_mid_wrlg <- (mean_all_min_wrlg+mean_all_max_wrlg)/2
mean_all_mid_clusters <- (mean_all_min_clusters+mean_all_max_clusters)/2

clean_wrlg <- ggpubr::ggscatter(data = mean_all,
          x = "mean_raw",
          y = "mean_wrlg",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
  geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(1, 250),
    breaks = c(mean_all_min_raw, mean_all_mid_raw, mean_all_max_raw), 
    labels = c(
  as.character(round(mean_all_min_raw, 1)),
  as.character(round(mean_all_mid_raw, 1)),
  as.character(round(mean_all_max_raw, 1))
               )
  )+
  scale_y_continuous(
    limits = c(1, 250),
    breaks = c(mean_all_min_wrlg, mean_all_mid_wrlg, mean_all_max_wrlg), 
    labels = c(
  as.character(round(mean_all_min_wrlg, 1)),
  as.character(round(mean_all_mid_wrlg, 1)),
  as.character(round(mean_all_max_wrlg, 1))
               )
  )+
  labs(x = "Raw dataset", y = "Clean dataset (Wear log)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Raw vs clean (clusters) scatterplot
clean_clusters <- ggpubr::ggscatter(data = mean_all,
          x = "mean_raw",
          y = "mean_clusters",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
   geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(1, 250),
    breaks = c(mean_all_min_raw, mean_all_mid_raw, mean_all_max_raw), 
    labels = c(
  as.character(round(mean_all_min_raw, 1)),
  as.character(round(mean_all_mid_raw, 1)),
  as.character(round(mean_all_max_raw, 1))
               )
  )+
   scale_y_continuous(
    limits = c(1, 250),
    breaks = c(mean_all_min_clusters, mean_all_mid_clusters, mean_all_max_clusters), 
    labels = c(
  as.character(round(mean_all_min_clusters, 1)),
  as.character(round(mean_all_mid_clusters, 1)),
  as.character(round(mean_all_max_clusters, 1))
               )
  )+
  labs(x = "Raw dataset", y = "Clean dataset (algorithm)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Histogram of deltas
## First, do some re-structuring
# Turning df into long form to perform computations
mean_all_long <- mean_all %>%
  pivot_longer(cols = c(delta_wrlg, delta_clusters),
                        names_to = "metric",
                        values_to = "delta") 


delta_hist <- ggplot(mean_all_long, aes(x = delta, fill = metric)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(x = "Delta", y = "Frequency") +
  scale_fill_manual(
    name = "Metric",
    values = c("#0072B2", "#D55E00"),
    labels = c("Algorithm dataset", "Wear log dataset")
    ) +
  scale_x_continuous(
    limits= c(-78, 78),
    breaks = c(-76, 0, 76),
    labels = c("-76", "0", "76")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        aspect.ratio = 1,
        plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm")) +
  ggpubr::rremove("legend")


# Combining the plots
## Make sure you have patchwork loaded in
p14 <- clean_wrlg + clean_clusters + delta_hist + 
  patchwork::plot_annotation("Average mEDI (daytime)",
  tag_levels = list(c("N1", "N2", "N3")),
  theme = theme(plot.title = element_text(hjust = 0.5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p14

```
# Calculating M10
```{r}
# Raw dataset
m10_raw_wday<- raw_dataset_wdays %>%
  group_by(Id, wDay) %>%
  summarize(bright_dark_period(log_zero_inflated(MEDI),
                               Datetime,
                               as.df = TRUE,
                               na.rm = TRUE),
            .groups = "drop"
  )

# Calculate mean across days   
m10_raw <- m10_raw_wday %>%
  group_by(Id) %>%
  mutate(brightest_10h_mean = exp_zero_inflated(brightest_10h_mean)) %>%
  summarize(m10_raw = mean(brightest_10h_mean, na.rm = TRUE),
            .groups = "drop")


# Wear log dataset
m10_wrlg_wday <- nw_wrlg_wdays %>%
  group_by(Id, wDay) %>%
  summarize(
      bright_dark_period(log_zero_inflated(MEDI),
                         Datetime,
                         as.df = TRUE,
                         na.rm = TRUE),
      .groups = "drop"
  )

# Calculate mean across days
m10_wrlg <- m10_wrlg_wday %>%
  group_by(Id) %>%
  mutate(brightest_10h_mean = exp_zero_inflated(brightest_10h_mean)) %>%
  summarize(
    m10_wrlg = mean(brightest_10h_mean, na.rm = TRUE),
    .groups = "drop"
  )

# Algorithm dataset
m10_clusters_wday <- nw_clusters_wdays %>%
  group_by(Id, wDay) %>%
  summarize(bright_dark_period(log_zero_inflated(MEDI),
                              Datetime,
                              as.df = TRUE,
                              na.rm = TRUE),
            .groups = "drop"
  )

# Calculate mean across days
m10_clusters <- m10_clusters_wday %>%
  group_by(Id) %>%
  mutate(brightest_10h_mean = exp_zero_inflated(brightest_10h_mean)) %>%
  summarize(
    m10_clusters = mean(brightest_10h_mean, na.rm = TRUE),
    .groups = "drop"
  )

# Combining the dfs 
temp <- full_join(m10_raw, m10_wrlg, by = "Id") #first, combine raw and wrlg data sets
m10_all <- full_join(temp, m10_clusters, by = "Id") #second, add clusters data set

# Calculating the delta
m10_all <- m10_all %>%
  mutate(delta_wrlg = m10_wrlg - m10_raw,
         delta_clusters = m10_clusters - m10_raw)
```

# Visualising M10
```{r}
# Raw vs clean (wrlg) scatterplot
clean_wrlg <- ggpubr::ggscatter(data = m10_all,
          x = "m10_raw",
          y = "m10_wrlg",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
  geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(10, 690),
    breaks = c(10, 450, 690), 
    labels = c("10", "450", "690")
  ) +
  scale_y_continuous(
    limits = c(10, 690),
    breaks = c(10, 450, 690), 
    labels = c("10", "450", "690")
  ) +
  labs(x = "Raw dataset", y = "Clean dataset (Wear log)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Raw vs clean (clusters) scatterplot
clean_clusters <- ggpubr::ggscatter(data = m10_all,
          x = "m10_raw",
          y = "m10_clusters",
          color = "Id",
          ggtheme = theme_bw(),
          size = 1.5) +
   geom_abline(intercept = 0, slope = 1, colour = "darkgrey", linetype = "dashed") + # adding unity line
  ggplot2::scale_x_continuous(
    limits = c(10, 690),
    breaks = c(10, 450, 690), 
    labels = c("10", "450", "690")
  ) +
  scale_y_continuous(
    limits = c(10, 690),
    breaks = c(10, 450, 690), 
    labels = c("10", "450", "690")
  ) +
  labs(x = "Raw dataset", y = "Clean dataset (algorithm)") +
  coord_fixed(ratio = 1) +
  ggpubr::rremove("legend") +
  theme(plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm"))

# Histogram of deltas
## First, do some re-structuring
# Turning df into long form to perform computations
m10_all_long <- m10_all %>%
  pivot_longer(cols = c(delta_wrlg, delta_clusters),
                        names_to = "metric",
                        values_to = "delta") 


delta_hist <- ggplot(m10_all_long, aes(x = delta, fill = metric)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(x = "Delta", y = "Frequency") +
  scale_fill_manual(
    name = "Metric",
    values = c("#0072B2", "#D55E00"),
    labels = c("Algorithm dataset", "Wear log dataset")
    ) +
  scale_x_continuous(
    limits= c(-116, 116),
    breaks = c(-116, 0, 116),
    labels = c("-116", "0", "116")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), 
        aspect.ratio = 1,
        plot.margin = unit(c(t=0,r=0.1,b=0.2,l=0), "cm")) +
  ggpubr::rremove("legend")


# Combining the plots
## Make sure you have patchwork loaded in
p15 <- clean_wrlg + clean_clusters + delta_hist + 
  patchwork::plot_annotation("M10m",
  tag_levels = list(c("O1", "O2", "O3")),
  theme = theme(plot.title = element_text(hjust = 0.5, size = 16),
                plot.margin = unit(c(t=0.1,r=0.1,b=0.1,l=0.1), "cm")))

p15
```
### Combining this into one big multiplot
```{r}

# Create multiplots with the plots - we create two of them as one would not fit all 12 plots
plots1 <- ggpubr::ggarrange(p1,p2,p4, p5, p6, p7, p8,
                   align = "v",
                   nrow = 7)

plots2 <- ggpubr::ggarrange(p9, p10, p11, p12, p13, p14, p15,
                   align = "v",
                   nrow = 7)

```

#### Saving the multiplot
```{r}
ggsave(filename = "results_fig10_a.png",
       plot = plots1, 
       width = 10,
       height = 15.5,
       dpi = 600,
       bg = "white",
       path = here::here("outputs", "fig10"))

ggsave(filename = "results_fig10_b.png",
       plot = plots2, 
       width = 10,
       height = 15.5,
       dpi = 600,
       bg = "white",
       path = here::here("outputs", "fig10"))
```

