---
title: "sleepstates_check"
author: "Carolina Guidolin"
date: "2024-05-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Perform a quality check for raw sleep diary files sanity check between wear log detected intervals and sleep and wake time from sleep diary 
As part of the non-wear analysis, we visualised time spent in each of the non-wear states: off (device not worn), on (device worn) and off for sleep (device not worn during sleep). As a sanity check, the time that the device is not worn during sleep should be similar to the time that the participants reported actually sleeping, as indicated by the sleep diary. For this reason, we follow a similar approach to the code above to visualise the amount of time participants spend asleep or awake according to the sleep diary. 

### NOTE! This script assumes you have run the script 02_wrlg_plots.Rmd.

```{r}
#Locate the sleep diary files
path_sleep <- here::here("00_raw_data", "sleepdiary")

# Get the files names from directory
sleepdiary = list.files(path_sleep, pattern="*.csv", full.names = TRUE)


#Using LightLogR, we import the files and label each timestamp as wake or sleep
sleepdiary_df <- LightLogR::import_Statechanges(sleepdiary,
                                                Datetime.format = "dmyHM",
                                                State.colnames = c("sleep", "out_ofbed"),
                                                State.encoding = c("sleep", "wake"),
                                                Id.colname = record_id,
                                                sep = ";",
                                                dec = ",",
                                                tz = tz)

```

## Quality check of the raw files
Here, we identify consecutive identidical States in the sleep diary
```{r}
error_messages_datetime <- character(0) # We create an empty vector where we can store the errors

# Group by ID and then apply the quality check within each group
sleepdiary_df %>%
  group_by(Id) %>%
  arrange(Datetime) %>%  # Ensure data is sorted by Datetime
  mutate(PrevState = lag(State)) %>%
  filter(State == PrevState) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(error_message = paste("Consecutive values at record_id:", Id, "datetime:", Datetime, "state:", State)) %>%
  pull(error_message) -> error_messages_states

# Check if there are error messages
if (length(error_messages_states) > 0) {
  # There are errors; print the error messages
  for (error_message in error_messages_states) {
    cat("Error:", error_message, "\n")
  }
} else {
  cat("No consecutive value errors found.")
}

```

```{r}
#This code led to the following errors: 

#Error: Consecutive values at record_id: 202 datetime: 2023-08-16 00:45:00 state: sleep 
#Error: Consecutive values at record_id: 202 datetime: 2023-08-17 06:15:00 state: wake 
#Error: Consecutive values at record_id: 205 datetime: 2023-08-31 09:47:00 state: wake 
#Error: Consecutive values at record_id: 205 datetime: 2023-08-31 23:15:00 state: sleep 
#Error: Consecutive values at record_id: 212 datetime: 2023-09-16 06:32:00 state: wake 
#Error: Consecutive values at record_id: 212 datetime: 2023-09-17 22:20:00 state: sleep 

#These were corrected directly on the csv file as they all stemmed from the participant accidentally logging the wrong data when reporting sleep and bed dates from the previous day. 
```

## Sleep diary interval transformation
Similar to what was done for the wearlog intervals, we now turn the state changes of the sleep diary to an interval
```{r}
sleep_int <- sleepdiary_df %>%
  LightLogR::sc2interval() %>%
  filter(!is.na(State))
```

#Plotting 
As part of the sanity check, we want to plot the distribution of sleep and wake for visualisation.
```{r}
#First, while all participants took part in the experiment for 7 days, they started at different times, and so it is important not just to approximate to 7 days (168 hours), but to calculate the specific 100% of "participation time" for each participant. For this, we first need to calculate the n of hours that each participant spends in the experiment.
exp_duration <- filtered_time %>%
  mutate(id_int = lubridate::interval(start, end),
         id_duration = lubridate::int_length(id_int) %>% as_hms()) %>% #tot hours of participation for each participant
  select(Id, id_duration)

#Calculate the amount of time spent in each state (either awake or alseep), for each participant across the whole experiment duration
int_duration_sleep <- sleep_int %>%
#calculate length of each interval
  mutate(int_length = int_length(Interval) %>% as_hms()) %>% 
  #filter out the last "awake" interval, since it goes to midnight of the final exp day, which is not actual data
  filter(row_number() != n()) %>% #this filters out the rows where row n is equal to tot number of rows, i.e. last enrty
  group_by(Id, State, .add = TRUE) %>%
  summarize(tot_intlength = sum(int_length) %>% as_hms()) %>%
  ungroup() 

#Loading in a package to create raincloud plots
library(ggdist)

# Normalise it to the total time they participated in the experiment
int_duration_sleep_clean <- int_duration_sleep %>%
  left_join(exp_duration, by ="Id") %>% #append the column containing total participation hours from the exp_duration df
  #note that exp_duration has been calculate earlier and can be reused here 
  mutate(id_duration_n= as.numeric(tot_intlength)/as.numeric(id_duration), #calculate percentage of time) #convert to numeric as division betwen difftime objects is not supported
         id_duration_per = id_duration_n*100) 

#Plot
sleep_wake_dist <- ggplot(data = int_duration_sleep_clean, aes(x=id_duration_per, y = State, fill = State)) +
  xlim(0,100) +
  ggdist::stat_halfeye(
    aes(colour = State,
        fill = State),
    adjust = .5,
    justification = -.25,
    point_color = NA,
    interval_color = NA
    ) +
  geom_boxplot(
    aes(color = State),
    width = .2,
    alpha = .2,
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(colour = State),
    fill = "white",
    height = .1,
    alpha = .3,
    size = 1.2
  ) +
  scale_fill_manual(
    values = c("sleep" = "darkgoldenrod2", "wake" = "darkcyan")) +
  scale_color_manual(
    values = c("sleep" = "darkgoldenrod2", "wake" = "darkcyan")) +
  scale_y_discrete(labels = c("sleep" = "Sleep", "wake" = "Wake")) +
  coord_fixed(ratio = 12) + # reduce white space between categorical variables on y axis
  theme_ggdist() +
  ggpubr::rremove("ylab") +
  ggpubr::rremove("y.ticks") +
  labs(title = "Sleep and wake distribution across the week", x = "Percentage of time (%)") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 24),
        axis.text.x = element_text(size = 16),
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16))

int_duration_sleep_clean$tot_intlength_hours <- as.numeric(int_duration_sleep_clean$tot_intlength) / 3600
  
second_x_axis_sleep <- ggplot(data = int_duration_sleep_clean, aes(x = tot_intlength_hours, y = 0.01)) +
  geom_blank() +  # Create an empty plot
  scale_x_continuous(limits = c(0, 166), breaks = c(0, 20, 40, 60, 80, 100, 120, 140, 166)) +
  coord_fixed(ratio = 4) +
  theme_ggdist() +
  labs(x = "Time (hours)")  +
  ggpubr::rremove("y.axis") +
  ggpubr::rremove("ylab") + 
  ggpubr::rremove("y.ticks") +
  ggpubr::rremove("y.text") +
  theme(axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 16)) 

# Combine the plots
sleep_dist_2axis <- sleep_wake_dist / second_x_axis_sleep 

print(sleep_dist_2axis)

```

## Save plots above
```{r}
ggsave(filename = "sleep_wake_dist.png",
       plot = sleep_wake_dist ,
       width = 8,
       height = 5,
       dpi = 600,
       path= here::here("outputs")
)

ggsave(filename = "sleep_dist_2axis.png",
       plot = sleep_dist_2axis,
       width = 8,
       height = 5,
       dpi = 600,
       path= here::here("outputs")
)
```


## Differences in sleep and wake time between sleep diary and wear log
We compare sleep times and wake times between these two sources.
```{r}
# Filter only sleep times in the wear log dataset 
sleepdiary_sleep <- sleepdiary_df %>%
  filter(State == "sleep")

# Check that each ID has only one sleep timestamp for each day
sleepdiary_sleep_clean <- sleepdiary_sleep %>%
  mutate(date = as.Date(Datetime,
                        tz = "Europe/Berlin")
         ) %>%
  group_by(Id, date) %>%
  summarise(n = n()) %>%
  filter(n>1)
## This leads to 17 instances with more than one sleep entries for each day. After visual check, we confirm that all instances are caused by the participant going to bed after midnight on one day and before midnight on the same day 

# Re-assign name
sleepdiary_sleep_clean <- sleepdiary_sleep

# Repeat the same procedure for wear log sleep entries
wrlgstates_sleep <- wrlg_states %>%
  filter(State == "sleep")

# Check for multiple entries on same day
wrlgstates_sleep_double <- wlrgstates_sleep %>%
  mutate(date = as.Date(Datetime,
                        tz = "Europe/Berlin")
         ) %>%
  group_by(Id, date) %>%
  summarise(n = n()) %>%
  filter(n>1)

## This leads to 23 instances, and after visual check most of them, as for the sleep diary, are caused by the participant going to bed after midnight on one day and before midnight on the same day. There are however also instanced of people who logged naps as sleep times (identified as naps using visual inspection): PID 206 and 231. Since we are interested in the main sleep period, we will remove these instances. 

nap_times <- data.frame(
  Id = c("206",
         "206",
         "206", 
         "206",
         "231",
         "231",
         "231"),
  Datetime = as.POSIXct(
    c(
      "2023-08-28 17:26:00",
      "2023-08-29 17:12:00",
      "2023-08-31 19:33:00",
      "2023-09-02 18:21:00",
      "2023-11-07 17:05:00",
      "2023-11-09 08:55:00",
      "2023-11-10 18:03:00"
      ),
    tz = tz)
  )

# Clean dataset from nap times         
wrlgstates_sleep_clean <- wrlgstates_sleep %>%
  anti_join(nap_times, by = c("Id", "Datetime"))

# Convert to data table to allow later join based on nearest timestamp
data.table::setDT(sleepdiary_sleep_clean)[, join_date := Datetime] #convert to data.table and append original column that we want to keep
data.table::setDT(wrlgstates_sleep_clean)[, join_date := Datetime] #convert to data.table and append original column that we want to keep

  #Find the nearest sleep onsets between the two datasets 
  sleep_onset_combined <- sleepdiary_sleep_clean[wrlgstates_sleep_clean, on = .(Id, join_date), roll = "nearest"] %>%
  #Perform a left join of the two datasets, for each row of wrlgstates_sleep (wear log entries) the nearest sleep onset timestamp is found
  #join_dat col is now the wearlog timestamp and Datetime is the sleep timestamp
  .[, tmp_diff := as.numeric(difftime(Datetime, join_date, units = "mins"))] 
  
# Visualise plot
p <- ggplot(sleep_onset_combined, aes(x = abs(tmp_diff))) +
  geom_histogram(binwidth = 3, fill = "steelblue", color = "white") +
  labs(
    title = "Time difference between Wear log reported removal of glasses \nbefore sleep and sleep time according to sleep diary",
    x = "Absolute difference (minutes)",
    y = "Count"
  ) +
  scale_x_continuous(
    breaks = c(0, 30, 60, 90, 120, 150, 180, 210, 240, 270, 300, 330)
  ) +
  theme_minimal() 

# Save plot 
ggsave(filename = "sleeponset_difference.png",
       plot = p,
       width = 6.5,
       height = 4,
       dpi = 600,
       bg = "white",
       path= here::here("outputs")
       )

print(p)

```
Two remarkable outliers:
- Participant 222 on 2023-10-29 reports going to bed at 23:53, but trying to go to sleep at 02:27:00 on the sleep diary. However, on the wear log, they report removing the glasses before going to sleep at 23:01. 
- Participant 230 on 2023-11-12 reports going to bed at 5:15 and also to sleep at 5:15 -- however takes off the glasses way earlier on the previous night, i.e. 23:32. 

