---
title: "blackbag_use"
author: "Carolina Guidolin"
date: "2024-05-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Note
This script should be run after 01_import and 02_datapreparation scripts. We make use of the save rds file to load in the clean dataset.

## Use of black bag (low illuminance values) to detect non-wear time
Aim: Here we analyse data from black bag use (low illuminance [mEDI] values). The first thing that we want to do is a visual inspection of the data to check how light and activity compare, for each State of the Wear log. Low activity is also often used as a way to understand non-wear time in field studies. Since we have both information (light and activity), we want to know how these two source compare to each other. 

```{r}
clean.LL.wrlg <- readRDS("clean_data.rds")
```


```{r}
#We will be plotting using log10 transformation, so we cannot have 0 values in our data set. For this reason, we add 1 to each value in MEDI and PIM

vis.LL.wrlg <- dataset.LL.wrlg

vis.LL.wrlg$MEDI <- vis.LL.wrlg$MEDI + 1
vis.LL.wrlg$PIM <- vis.LL.wrlg$PIM + 1

#To be able to visually observe how light and activity relate to each other in our data set, we cannot plot all our data points (too many). So, we take median values of each Id, for each State (off/on/sleep)

median.LL.wrlg <- vis.LL.wrlg %>%
#  filter(!State == "sleep") %>%
  group_by(Id, State) %>%
  summarise(median_medi = median(log10(MEDI), na.rm = TRUE),
            median_pim = median(log10(PIM), na.rm = TRUE),
            qr1_medi = quantile(log10(MEDI), 1/4, na.rm = TRUE), #first quantile
            qr1_pim = quantile(log10(PIM), 1/4, na.rm = TRUE), #first quantile
            qr3_medi = quantile(log10(MEDI), 3/4, na.rm = TRUE), #third quantile
            qr3_pim = quantile(log10(PIM), 3/4, na.rm = TRUE)) %>% #third quantile
  ungroup()

#Let's create a simple scatterplot where median mEDI and PIM are plotted against each other, including error bars
median_scatter <- 
  ggplot(median.LL.wrlg, aes(x = median_medi, y = median_pim, color = State, fill = State)) +
  geom_point() +
    #position = position_jitter(width = 0.5, height = 0.5)) 
 # xlim(-5, 2.5) +
 # ylim(0, 750) +
  scale_fill_manual(
    values = c("off" = "darkred", "on" = "#0072B2" 
               #"sleep" = "darkgoldenrod2"
               )) +
  scale_color_manual(
    values = c("off" = "darkred", "on" = "#0072B2"
               #"sleep" = "darkgoldenrod2"
               )) +
  labs(title = "Relationship between light and activity", x = "mEDI", y = "PIM") +
  geom_errorbar(aes(ymin = qr1_pim, ymax = qr3_pim)) +
  geom_errorbarh(aes(xmin = qr1_medi, xmax = qr3_medi)) +
  theme_minimal() +
  scale_x_continuous(transform = "symlog")

#For the states sleep and off, there are a lot of overlapping points, which is why only a few points are shows. While this plot shows that we have a lot of variability, especially in PIM, we can't really say much about the individual datapoints. 

```

## Improving visualisation above
To improve the visualisation above, we can create a simple scatterplot without error bars, and add boxplots as marginal plots.
```{r}

set.seed(1) #we will use geom_jitter, which uses random sampling. For reproducibility we need to set.seed(1) so that running the script multiple times will lead to the same plot

scatter <- ggplot(median.LL.wrlg, aes(x = median_medi, y = median_pim)) + 
  geom_jitter(aes(fill = State, color = State), size = 2.5, alpha = 0.6,
              position = position_jitter(width = .3, height = .3)) + #adjust position of points to avoid overlap
   scale_fill_manual(
    name = "Wear log \n'state'",
    values = c("off" = "darkred", "on" = "#0072B2", "sleep" = "darkgoldenrod2"),
    labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  scale_color_manual(
    name = "Wear log \n'state'",
    values = c("off" = "darkred", "on" = "#0072B2", "sleep" = "darkgoldenrod2"),
    labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  scale_x_continuous(
    breaks = c(-1, 1,3), 
    labels = c(expression(10^-1), expression(10^1), expression(10^3)), #ensure that the log expression is labelled correctly 
    limits = c(-1, 4)
  ) +
  scale_y_continuous(
    breaks = c(-1, 1, 3),
    labels = c(expression(10^-1), expression(10^1), expression(10^3)), #ensure that the log expression is labelled correctly 
    limits = c(-1, 4)
  ) +
  labs(x = "Median of log illuminance (mEDI, lx)",
       y = "Median of log activity (PIM)") +
  coord_fixed(ratio=1) + #make sure plot is a square
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5, size = 18),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12, hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.position = "bottom") +
  guides(fill = guide_legend(nrow = 1))
  
#Add a boxpot as a marginal plot using package ggExtra
library(ggExtra)
library(ggpubr)

marginal <- ggExtra::ggMarginal(scatter, type = "boxplot", groupColour = TRUE, groupFill = TRUE)


plot_marginal <- ggpubr::annotate_figure(marginal,
                                 #Add title
                                 top = text_grob("Relationship between illuminance \nand activity across wear log states",
                                                            color = "black",
                                                            face = "plain",
                                                            size = 18))
# Saving the plot
ggsave(filename = "results_fig04a.svg",
       plot = plot_marginal ,
       width = 5,
       height = 5.5,
       dpi = 600,
       bg = "white",
       path= "H:/nonwear_detection/preprint_figures/results/fig04")

```

#Hexagonal plots across states for all participants
This allows us to compare light and activity across the whole week - this can be done for on, off and sleep states alike
```{r}
##List of all unique participant Ids
par_ids <- unique(dataset.LL.wrlg$Id)

for (participant in par_ids) {
  subset_df = dataset.LL.wrlg %>%
    filter(Id == participant) %>%
    filter(State == "on") %>%
    select(Id, State, PIM, MEDI)
  
  plot <- ggplot(data = subset_df, 
          aes(x = MEDI, y = PIM)) +
    geom_hex(bins = 74)  +
    scale_fill_gradient(low = "lightgreen", high = "darkgreen") +
    scale_y_continuous(trans = "symlog",
                     breaks = c(-10^(5:0), 0, 10^(0:5)),
                     labels = function(x) format(x, scientific = FALSE, big.mark = " ")) +
    scale_x_continuous(trans = "symlog",
                     breaks = c(-10^(5:0), 0, 10^(0:5)),
                     labels = function(x) format(x, scientific = FALSE, big.mark = " ")) +
    labs(title = paste0(participant, " light and activity")) +
 #   coord_fixed(ratio = 1) +
    theme_bw()
  
  filename <- paste0("p", participant, "_pim_medi.png")
  
  ggsave(filename = filename,
         plot = plot,
         width = 8,
         height = 8,
         dpi = 600,
         bg = "white",
         path= "G:/cyepi/code/outputs/light_activity_prc/hexagonalplots/on")
    
} 
```



