---
title: "alg_performance"
author: "Carolina Guidolin"
date: "2024-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualisation of where the low mEDI cluster detection algorithm goes wrong 

### First, we identify clusters of low mEDI using the input parameters known to perform best:
- Low illuminance threshold: 1 mEDI lux 
- Minimum length of the cluster: 10 minutes 
- Maximum interruption within cluster: 0 minutes

```{r}

#Empty list to store classification results
prc_list <- list()

classified_prc_medi <- alg_perf(
    dataset = dataset.LL.wrlg,
    low_var = "MEDI",
    min_length = 60, #10 minutes (n of observations)
    max_interrupt = 0, #0 minutes (n of observations)
    threshold = 1 #1 mEDI lux
  )

```

### Second, we need to create intervals for each classification length 
```{r}
medi_sc <- classified_prc_medi %>%
 group_by(Id) %>%
 mutate(interval_group = consecutive_id(classification)) %>% #need to set default for first value 
  group_by(Id, interval_group) %>%
summarise(classification = first(classification), #keep only first value for each Id and interval_group
            xmin = first(Datetime),
            xmax = last(Datetime)) %>%
  ungroup() 

```

### Third, and last, we need to visualise the classification as a time series, i.e. every day of the week, for each participant, with shaded areas for each classification type (TP, TN, FP, FN) and with the line colour of mEDI coding for the Wear log State.
```{r}
##List of all unique participant IDs
participant_ids <- unique(dataset.LL.wrlg$Id)

##Create a for loop to generate a plot for each participant 
for (selected_id in participant_ids) {
  
 subset_df <- dataset.LL.wrlg %>%
  filter(Id == selected_id)


##Calculate the limits of the shaded areas according to the Wear Log
alg.times <- medi_sc %>%
  filter(Id == selected_id) %>%
  mutate(xmax = (xmax + lubridate::seconds(10)), #need to add 10 seconds to match next start time
         Day.data.min = xmin %>% format(format = "%a"), #lower limit of the shaded area, needed for geom_rect
         Day.data.max = xmax %>% format(format = "%a"), #upper limit of shaded area, needed for geom_rect
         xmin = xmin %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         xmax = xmax %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         Day.data.min = factor(Day.data.min, levels = unique(Day.data.min)),
         Day.data.max = factor(Day.data.max, levels = unique(Day.data.max)))

##Some intervals on the Wear log that we want to plot cross day boundaries. We thus need to split them in 2 intervals, since we plot each day on a different facet of the plot 

#Step 1: collect the part of the data that does not cross day boundaries
alg.times1 <-
  alg.times %>%
  filter(Day.data.min == Day.data.max)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
alg.times2 <-
alg.times %>% filter(Day.data.min != Day.data.max) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        xmin=c(xmin, hms::as_hms("00:00:00")),
        xmax=c(hms::as_hms("23:59:59"), xmax),
        Day.min=c(Day.data.min, Day.data.max),
        Day.max=c(Day.data.min, Day.data.max)
      )
    )
  ) %>%
  select(classification, new_entry) %>%
  unnest(new_entry)

#Step 3: Combine the two again in a single df
alg.combined <- 
  bind_rows(alg.times1 %>% rename(Day.min = Day.data.min),
            alg.times2) %>%
  select(-Day.max) %>%
  rename(Day.data = Day.min) 

alg.combined$classification <- factor(alg.combined$classification, levels = c("TP", "TN", "FP", "FN", "NA")) #turn State into factor, NA represents the sleep period here


p <- gg_day(data = subset_df, #base plot defaults to plotting mEDI on y axis
            linewidth = 0.05, 
            aes_col = State, 
            geom = "line",
            group = date(Datetime),
            format.day = "%a",
            x.axis.label = "Time", 
            y.axis.label = "Illuminance (mEDI, lx)",
            title = "Algorithm performance for low mEDI clusters",
            subtitle = "mEDI threshold = 1 lux, min. length = 10 minutes, max. interruption = 0 minutes") +
  geom_rect(data=alg.combined, aes(xmin= xmin, xmax = xmax, ymin =0, ymax = Inf, fill = classification), #adding classification output
                    inherit.aes = FALSE,
                    alpha = 0.15) +
    scale_fill_manual(
     name = "Algorithm classification",
     values = c("TP" = "darkgreen",
                "TN" = "#0072B2",
                "FP" = "darkred",
                "FN" = "darkgoldenrod2"),
                na.value = "white") +
  scale_color_manual(
    name = "Wear log 'state'",
    values = c("on" = "#0072B2", "off" = "darkred", "sleep" = "black"),
    labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
   guides(
    color = guide_legend(override.aes = list(linewidth = 2)) # Make lines thicker in Wear log legend
  ) +
  guides(linetype = guide_legend(override.aes = list(size = 4))) +
  theme(plot.title = element_text(hjust = 0.5, size = 24, face = "plain"),
        plot.subtitle = element_text(hjust = 0.5, size = 16),
        axis.text.x = element_text(size = 16),
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14))

filename <- paste0("p", selected_id, "alg_perf_mEDI.png")
ggsave(filename = filename,
       plot = p,
       width = 10,
       height = 14,
       dpi = 600,
       bg = "white",
       path= "H:/nonwear_detection/preprint_figures/alg_class")
}
```

## Can we extract information about the misclassified non-wear intervals?
We would like to identify patterns in the false negative and false positive intervals, to understand what leads to misclassifications of these specific intervals. We want to collect information about this in a table:
- Was the bag used, according to what the participant reported?
- What was the average mEDI of these periods?
- How long were these intervals?

### First, we want to identify these intervals
```{r}
classification_int <- classified_prc_medi %>%
 group_by(Id) %>%
 mutate(interval_group = consecutive_id(classification)) %>% #need to set default for first value 
  group_by(Id, interval_group) %>%
summarise(classification = first(classification), 
          bag = first(bag), 
          is_low_medi_cluster = first(is_low_medi_cluster),
          mean_MEDI = mean(MEDI), #keep only first value for each Id and interval_group
          xmin = first(Datetime),
          xmax = last(Datetime), 
          class_duration = difftime(xmax, xmin, tz= tz, units = "mins")) %>%
  ungroup() %>%
  filter(classification == "FN" | classification == "FP") 
  
```

### Identifying reasons why these intervals were misclassified
Given the misclassified interval (false negative, FN, and false positive, FP), the mean MEDI of that interval, and its length (lower or higher than 3 minutes), we will identify the following reasons leading to misclassification:

#### False negatives (FN) intervals
1. If the bag was reportedly not used, and the mean MEDI > 1 lux, then the interval was misclassified because the bag was not used and so no clusters of low illuminance could be identified.
2. If the bag was reportedly used, and mean MEDI > 1 lux, and interval length <= 3 minutes, then this is a transition state
3. If the bag was reportedly used, and mean MEDI > 1 lux, and interval length > 3 minutes, then this is either a limitation of the algorithm (the non-wear interval was lower than 10 minutes), or the bag was actually not used
4. If the bag was reportedly used, and mean MEDI  <= 1 lux, and interval length <= 3 minutes, then this is a transition state

#### False positives (FP) intervals
1. If mean MEDI <= 1 lux and interval length <= 3 minutes, then this is a transition state
2. If mean MEDI <= 1 lux and interval length > 3 minutes, this is a low mdi interval which is hence detected as a cluster
```{r}
# Setting a lux threshold for mEDI
threshold = 1

# Setting a length threshold for what we define a transition state
transition_cutoff = as.difftime(3, unit = "mins")

misclass_reasons <- classification_int %>%
  mutate(
    reason = case_when(
      classification == "FN" & bag == 0 & mean_MEDI > threshold ~ "bag not used, no clusters identified",
      classification == "FN" & bag == 1 & mean_MEDI > threshold & class_duration <= transition_cutoff ~ "transition state",
      classification == "FN" & bag == 1 & mean_MEDI > threshold & class_duration > transition_cutoff ~ "algorithm limitation or bag actually not used",
      classification == "FN" & bag == 1 & mean_MEDI <= threshold & class_duration <= transition_cutoff ~ "transition state",
      classification == "FP" & mean_MEDI <= threshold & class_duration <= transition_cutoff ~ "transition state",
      classification == "FP" & mean_MEDI <= threshold & class_duration > transition_cutoff ~ "low medi and cluster detected",
      TRUE ~ "visual_check" # Fallback case
    )
  )

#Compiling into a summary table
classification_summary <- misclass_reasons %>%
  group_by(classification, reason) %>%
  summarise(count = n()) 
```
#### Performing visual checks 
The classification above led to 4 instances, i.e. intervals, which could no be assigned to any misclassification reasons we reported. Hence, we perform a visual check here to see why these instances were misclassified.
```{r}
# Identify the instances 
vis_chek <- misclass_reasons %>%
  filter(reason == "visual_check")

# Case 1: The FN intPID 212 from 2023-09-11 21:13:00 to 2023-09-11 21:21:50 reported using the bag, but the mean MEDI is < 1 lux. The interval duration is 9 minutes. From visual inspection, we can see that they did use the bag. Hence, this misclassification is likely due to an algorithm limitation of detecting clusters > 10 minutes. We change this manually. 

misclass_reasons_1 <- misclass_reasons %>%
  mutate(reason = ifelse(Id == "212" & xmax == as.POSIXct("2023-09-11 21:21:50", tz = tz), "algorithm limitation or bag actually not used", reason))

# Case 2: PID 222 reported not using the bag from 2023-10-26 16:04:02 to 2023-10-26 16:31:52 and mean MEDI <1 lux in this 28 minute interval. Upon visual inspection, we identify that this was not detected as a low illuminance cluster because there are multiple timepoints where MEDI is above threshold (1 lux), even if the average is below threshold. We change this manually and label is as algorithm limitation. 

misclass_reasons_2 <- misclass_reasons_1 %>%
  mutate(reason = ifelse(Id == "222" & xmax == as.POSIXct("2023-10-26 16:31:52", tz = tz), "algorithm limitation or bag actually not used", reason))

# Case 3: PID 225 reported using the bag from 2023-10-31 05:42:05 to 2023-10-31 05:45:55 (4 minute FN interval). The mean MEDI here is also below threshold. This is an algorithm limitation, as the interval is too short to be detected by our cluster function. We can label it as such manually. 

misclass_reasons_3 <- misclass_reasons_2 %>%
  mutate(reason = ifelse(Id == "225" & xmax == as.POSIXct("2023-10-31 05:45:55", tz = tz), "algorithm limitation or bag actually not used", reason))

# Case 4: This FN interval for PID 230 lasts between 2023-11-09 10:56:06 and 2023-11-09 11:04:56. The participant reported using the bag and mean MEDI is below threshold. However, by visual inspection, we identify several instances where the mean MEDI was above threshold. Since we allow for no interruptions in our detection, this leads to misclassification. We can change this manually 

misclass_reasons_4 <- misclass_reasons_3 %>%
  mutate(reason = ifelse(Id == "230" & xmax == as.POSIXct("2023-11-09 11:04:56", tz = tz), "algorithm limitation or bag actually not used", reason))

```

#### Recoding the reasons for misclassification after visual checks
We would like to update the classification_summary table based on the changes made to the 4 instances above.
```{r}
# Load janitor package for data cleaning
library(janitor)

# First, let's rename the data frame with a better name
misclass_reasons_clean <- misclass_reasons_4

# Then, we create the summary table
misclass_summary <- misclass_reasons_clean %>%
  group_by(reason) %>%
  summarise(count = n()) %>%
  mutate(per_instances = (count/sum(count))*100,
         per_instances = round(per_instances, digits = 3)) %>%
  #Adding a row with tot numbers
  janitor::adorn_totals(where = c("row"))

# As expected and desired, this table has no visual_check instances and the number of reason = "algorithm limitation or bag actually not used" has increased by 4 compared to the table classification_summary
```

## Compiling the reasons for misclassification in a table 
We want to create a table that can be incorporated in the pre-print where the various reasons for misclassification are illustrated, along with the reason we identified leading to the misclassification.
```{r}
# Load gt package to create a publication-ready table
library(gt)

misclass_table <- misclass_summary %>%
  dplyr::arrange(count) %>% # Arranging in ascending order
  gt() %>% # Turning this into a gt object for further editing of the table 
  gt::tab_header(title = md("**Instances of algorithm misclassification**")) %>%
  gt::tab_stubhead(label = "reason") %>%
  #Re-labelling the columns
  gt::cols_label(reason = "Instance",
             count = "N",
             per_instances = "N (%)") %>%
  #Re-labelling row names
  gt::text_case_match("bag not used or transition state" ~ "Bag not used or transition state",
                      "algorithm limitation or bag actually not used" ~ "Algorithm limitation or bag not actually used",
                      "bag not used, no clusters identified" ~ "Bag not used",
                      "low medi and cluster detected" ~ "Low illuminance during wear",
                      "transition state" ~ "Transition state")
```

#### Saving the table
```{r}
# Load chromote package needed to save using gt
library(chromote)

# Make sure you set your path to Chrome using Sys.setenv(CHROMOTE_CHROME = "path/to/chrome.exe") and check this was correct by running chromote::find_chrome()

gt::gtsave(misclass_table,
           filename = "results_fig06.png",
           path = "H:/nonwear_detection/preprint_figures/results")
```
## Adding a padding to the transition states 
We have now identified that a large part of the misclassified intervals are due to transition states. With this knowledge, we can add a padding to each of these intervals, so that is_low_medi_cluster would expand of 3 minutes on both ends, i.e. we are more liberal in what we define as transition state here. How do we implement the padding? There are a couple of steps to this:
1. The information on which intervals are transition states is currently stored in misclass_reasons_clean. On the other hand, the output of non-wear detection based on clusters of low illuminance is stored in classified_prc_medi. We need to label each Datetime of classified_prc_medi as transition state, where needed.

### Step 1: we run again the cluster detection on the dataset and call this nonwear_algorithm
```{r}

#Empty list to store classification results
prc_list <- list()

nonwear_algorithm <- alg_perf(
    dataset = dataset.LL.wrlg,
    low_var = "MEDI",
    min_length = 60, #10 minutes (n of observations)
    max_interrupt = 0, #0 minutes (n of observations)
    threshold = 1 #1 mEDI lux
  )
```

### Step 2: we join the transition states from the "interval df" (i.e. misclass_reasons_clean) to the "continuous" dataset (i.e. nonwear_algorithm)
For this, we use the package data.tables
```{r}
library(data.table)

# We turn our data frames into data tables
data.table::setDT(misclass_reasons_clean)
data.table::setDT(nonwear_algorithm)

# Set keys for efficient range joins. Key are columns that are indexed for fast identification when sorting (next step)
data.table::setkey(misclass_reasons_clean, xmin, xmax)

# We only want to move the reason = transition state over to the other dataframe (we do not care about the other reasons)
transitions <- misclass_reasons_clean %>%
   filter(reason == "transition state")

# Joining the two data.tables
nw_alg <- nonwear_algorithm[,
  reason := transitions[on = .(Id, xmin <= Datetime, xmax >= Datetime), # for each Datetime of nonwear_algorithm, it checks all rows in transitions where xmin <= Datetime and xmax >= Datetime (based on Id)
                        mult = "first", # if multiple intervals match a Datetime, the first matching reason is assigned
                        reason]] # the reason column is assigned to nonwear_algorithm

```

### Step 3: add transition pad to non-wear intervals
We want to use a function that performs the following operations:
1. Detect when is_low_medi_cluster changes from 1 to 0.
2. When such a change occurs, check if the previous row has reason == "transition state".
3. If true, find the start of the "transition state" sequence (i.e., the first occurrence of reason == "transition state" before this row).
4. Set is_low_medi_cluster = 0 for all rows from the start of the "transition state" to the row just before the detected change from 1 to 0.

We create this function in the file add_pad_fun.R, where each step is described. Here, we call the function and apply to the relevant dataset.
```{r}
# Apply function that adds padding to non-wear intervals, when they are preceded by a transition state
# Note that this will automaticall print all instances where the values of is_low_medi_cluster are modified
# If you want to change this, you have to do this in the function itself
nw_alg_padded <- add_transition_pad(nw_alg)
```

