---
title: "metrics_comparison"
author: "Carolina Guidolin"
date: "2024-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Calculating metrics using different datasets
The aim of this script is to calculate and compare light metrics on three different datasets:
1. A raw dataset, where non-wear information is ignored (i.e., we include all data regardless of whether it is wear or non-wear)
2. A dataset where non-wear information is identified using self-reported non-wear time (wear log)
3. A dataset where non-wear information is identified by detecting clusters of low illuminance

### Creation of the three datasets
Before performing the comparison, we need to identify these three datasets. 
```{r}
# 1. Raw dataset: this is the raw data, i.e. dataset.LL.wrlg
raw_dataset <- dataset.LL.wrlg

# 2. Non-wear by wear log dataset. We need to create this from dataset.LL.wrlg
nw_wrlg <- dataset.LL.wrlg %>% 
  mutate(MEDI = if_else(State == "off", NA, MEDI))

##Think about whether you need to change the function alg_perf and generate_prc for the sleep - since now all the sleep states are labelled as NA. Should this be done outside of the function?

```

