---
title: "prc_fig_combined"
author: "Carolina Guidolin"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aim
This script combines two plots: 
1) Scatterplot of the relationship between activity and light levels across Wear log states (see script bag_use_visualinspection)
2) Precision recall curve for low illuminance and low activity across all participants (prc_combined in script bag_activity_prc)

As this script only serves the purpose of generating a multiplot with these two plots, ensure that you have run the scripts with the individual plots prior to running the following chunks of code. 

### Combining the plots
```{r}
library(cowplot)


medi_pim_fig <- cowplot::plot_grid(plot_marginal,
                          prc_combined,
                          labels = c("A", "B"),
                          nrow = 1,
                          align = "v",
                          axis = "tblr",
                          rel_widths = c(1,1),
                          rel_heights = c(1,1))

# Save the plot
ggsave(filename = "results_fig04_all.svg",
       plot = medi_pim_fig,
       height = 5.5, 
       width = 10,
       dpi = 600,
       bg = "white",
       path= "H:/nonwear_detection/preprint_figures/results/fig04")
```

