---
title: "smoothed_prc"
author: "Carolina Guidolin"
date: "2024-09-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Note: this script has to be run after bag_activity_prc.Rmd

## What if we smoothed the data before applying the cluster detection algorithm 
1) Apply a log transformation to raw data
2) Calculate sliding median window 10 minutes 
3) Normalise to max per dataset 
4) Apply cluster detection algorithm (suggested parameters: 10min periods with 2min interruptions of activity < 0.1)

##Step 1: Apply log transformation
```{r}
#Before we transform into log, we replace 0 values with low numbers. We will do this on a dataframe called log.transf.LL.wrlg to avoid changing the df dataset.LL.wrlg directly

log.transf.LL.wrlg <- dataset.LL.wrlg

log.transf.LL.wrlg$MEDI[log.transf.LL.wrlg$MEDI == 0] <- 0.0001
log.transf.LL.wrlg$PIM[log.transf.LL.wrlg$PIM == 0] <- 0.0001

#Take log10 of both MEDI and PIM
log.LL.wrlg <- log.transf.LL.wrlg %>%
  select(Id, Datetime, PIM, MEDI, State, bag) %>%
  mutate(log.PIM = log10(PIM),
         log.MEDI = log10(MEDI))
```

##Step 2: Calculate sliding median window of 10 minutes
We will need the zoo package for this. Since our device measures light every 10 seconds, if we want to apply a 10 minutes sliding window, the number of observations will be 60 (10 minutes = 600 seconds, and 600seconds / 10 seconds/observations = 60 observations), which is what we need to specify in the width argument of rollapplyr.  
```{r}
library(zoo)

rolled.df <- log.LL.wrlg %>%
  mutate(rolled_median_MEDI = zoo::rollapplyr(log.MEDI, width = 60, FUN = median, fill = NA, align = "center"),
         rolled_median_PIM = zoo::rollapplyr(log.PIM, width = 60, FUN = median, fill = NA, align = "center"))

```

```{r}
medithresholds <- seq(1, 10, by = 1)
log_thr_medi <- log10(medithresholds)
max_interrupt = 0 
min_length = 60



#Empty list to store classification results
prc_list <- list()

for (threshold in log_thr_medi) {

prc_result <- generate_prc(
    dataset = rolled.df,
    low_var = "rolled_median_MEDI",
    min_length = 60, #10 minutes (n of observations)
    max_interrupt = 0, #0 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_m_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_list[[as.character(threshold)]] <- prc_result
  
}

#Turn list into df
prcurve_medi_sm <- bind_rows(prc_list)
#Turn threshold to numeric for plotting
prcurve_medi_sm$threshold <- as.numeric(prcurve_medi_sm$threshold)
  
```

```{r}
prc_all_ids <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  geom_point(data = prcurve_medi_sm, aes(x = TPR, y = PPV, colour = threshold)) +
   scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(log10(1), log10(10)),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Precision-recall curve for detection of \nnon-wear intervals based on illuminance thresholds",
       subtitle = "Min. length = 10 min, max. interruption = 0 min") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        plot.subtitle = element_text(hjust = 0.5, size = 14, colour = "grey33"),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 8),
        legend.key.size = unit(4, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal") +
  coord_fixed(ratio = 1) 
```




##Step 4a: Apply cluster detection for low activity
This chunk is the same as we have previously run for the non-log transformed raw data (see bag_activity_prc). 
```{r}
#Create a sequence of PIM thresholds at 5 unit steps
pimthresholds <- c(5,10, 15, 20, 25, 30, 35, 40, 45, 50)

log_thr_pim <- log10(pimthresholds)

#Parameters to feed into function
min_length = 60 # 10 minutes
max_interrupt = 6 # 1 minute

#Empty list to store classification results
prc_pim_list_sm <- list()
  
  for (threshold in log_thr_pim) {

  pim_clust_sm <- rolled.df %>%
    ungroup() %>%
    mutate(State = if_else(State == "sleep", "NA", State)) %>%  #converting all sleep values to NAs, since we do not want to use these for our classification 
   # filter(!is.na(rolled_median_PIM)) %>% #account for the start and end of the rolling median, which has NAs   
   # Replace `low_activity` by the (logical) variable for which the clusters 
   # should be found.
   mutate(low_pim = rolled_median_PIM < threshold) %>% 
   # Nest by groups, e.g., participant, period. Replace with grouping variables 
   # in your data, such that the function is applied separately per consecutive 
   # timeseries (e.g., from one subject during one period).
    nest_by(Id) %>% 
   mutate(
    # Replace `low_activity` by the (logical) variable for which the clusters 
    # should be found.
    data = list(
      data_find_clusters(
        data, 
        low_pim, 
        min_length = min_length, 
        max_interrupt = max_interrupt, 
        cluster_name = "low_pim_cluster"
      )
    )
  ) %>%
  unnest(cols = data) %>%
  ungroup() 


  pimclusters_clean_sm <- pim_clust_sm %>%
   select(Id, Datetime, State, bag, is_low_pim_cluster) %>%
    mutate(
      State = case_when(
      State == "on" ~ 1, 
      State == "off" ~ 0,
      is.na(State) ~ NA_real_), # keep sleep states as NA
    is_low_pim_cluster = case_when(
      is_low_pim_cluster == TRUE ~ 0,
      is_low_pim_cluster == FALSE ~ 1
    )) 

#We want to build a precision recall curve for our classifier algorithm, which, based o our "Ground truth", i.e. the wear log state values (on = 1 and off = 1), classifies the performance of our algorithm for detecting non-wear as follows:
#1) State = 0, cluster = 0 -> true positive
#2) State = 0, cluster = 1 -> false negative
#3) State = 1, cluster = 1 -> false positive
#4) State = 1, cluster = 0 -> true negative

  pimclusters_clean_sm <- pimclusters_clean_sm %>%
   mutate(classification = case_when(
      State == 0 & is_low_pim_cluster == 0 ~ "TP",
      State == 0 & is_low_pim_cluster == 1 ~ "FN",
      State == 1 & is_low_pim_cluster == 0 ~ "FP",
      State == 1 & is_low_pim_cluster == 1 ~ "TN",
     .default = NA_character_))

  prc_pim_sm <- pimclusters_clean_sm %>%
    group_by(classification) %>%
   summarise(count = n())  %>%
   pivot_wider(names_from = classification, values_from = count, values_fill = list(count=0)) %>%
    mutate(TPR = TP/(TP+FN), #this is the formula for TPR (true positive rate)
           FPR = FP/(FP+TN), #this is the formula for the FPR (false positive rate)
           PPV = TP/(TP+FP), #positive predictive value
           NPV = TN/(FN+TN), #negative predictive value
           threshold = threshold)  #adding manually which threshold I am considering here 

  #Add the result to the list
   prc_pim_list_sm[[as.character(threshold)]] <- prc_pim_sm
  }

#Turn the list with the results into a df 
prcurve_pim_sm<- bind_rows(prc_pim_list_sm)

#Turn thresholds from character to factor for plotting
prcurve_pim_sm$threshold <- as.numeric(prcurve_pim_sm$threshold)
```

#Step4b: Apply cluster detection for low mEDI
```{r}
#Define threholds for cluser detection algorithm
medi_threshold = log10(1) 
max_interrupt = 0 
min_length = 60

#Empty list to store classification results

medi_clusters_sm <- rolled.df %>%
  ungroup() %>%
  mutate(State = if_else(State == "sleep", NA_character_, State)) %>%  #converting all sleep values to NAs, since we do not want to use these for our classification 
  # filter(!is.na(rolled_median_MEDI)) %>% #account for the start and end of the rolling median, which has NAs
  # Replace `low_activity` by the (logical) variable for which the clusters 
  # should be found.
  mutate(low_medi = rolled_median_MEDI < medi_threshold) %>% 
  # Nest by groups, e.g., participant, period. Replace with grouping variables 
  # in your data, such that the function is applied separately per consecutive 
  # timeseries (e.g., from one subject during one period).
  nest_by(Id) %>% 
  mutate(
    # Replace `low_activity` by the (logical) variable for which the clusters 
    # should be found.
    data = list(
      data_find_clusters(
        data, 
        low_medi, 
        min_length = min_length, 
        max_interrupt = max_interrupt, 
        cluster_name = "low_medi_cluster"
      )
    )
  ) %>%
  unnest(cols = data) %>%
  ungroup() %>%
  select(Id, Datetime, State, MEDI, low_medi, is_low_medi_cluster)


mediclusters_clean_sm <- medi_clusters_sm %>%
  select(Id, Datetime, State, is_low_medi_cluster) %>%
  mutate(
    State = case_when(
      State == "on" ~ 1, 
      State == "off" ~ 0,
      is.na(State) ~ NA_real_), # keep sleep states as NA ),
    is_low_medi_cluster = case_when(
      is_low_medi_cluster == TRUE ~ 0,
      is_low_medi_cluster == FALSE ~ 1
    )) 

#We want to build a precision recall curve for our classifier algorithm, which, based o our "Ground truth", i.e. the wear log state values (on = 1 and off = 1), classifies the performance of our algorithm for detecting non-wear as follows:
#1) State = 0, cluster = 0 -> true positive
#2) State = 0, cluster = 1 -> false negative
#3) State = 1, cluster = 1 -> false positive
#4) State = 1, cluster = 0 -> true negative

mediclusters_clean_sm <- mediclusters_clean_sm %>%
  mutate(classification = case_when(
    State == 0 & is_low_medi_cluster == 0 ~ "TP",
    State == 0 & is_low_medi_cluster == 1 ~ "FN",
    State == 1 & is_low_medi_cluster == 0 ~ "FP",
    State == 1 & is_low_medi_cluster == 1 ~ "TN",
    .default = NA_character_))

prc_medi_sm <- mediclusters_clean_sm %>%
  group_by(Id, classification) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = classification, values_from = count, values_fill = list(count=0)) %>%
  mutate(TPR = TP/(TP+FN), #true positive rate
         FPR = FP/(FP+TN), #false positive rate
         PPV = TP/(TP+FP), #positive predictive value
         NPV = TN/(FN+TN), #negative predictive value
         threshold = threshold)  #adding manually which threshold I am considering here 

 #Add the result to the list
  prc_medi_list_sm[[as.character(threshold)]] <- prc_medi_sm
}

#Turn the list with the results into a df 
prcurve_medi_sm<- bind_rows(prc_medi_list_sm)

#Turn thresholds from character to factor for plotting
prcurve_medi_sm$threshold <- as.numeric(prcurve_medi_sm$threshold)
```

##Let's start to see how the cluster detection works on the group smoothed data, i.e. for all participants 
```{r}
prc_smoothed <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  #First plot: activity
  geom_point(data = prcurve_pim_sm, aes(x = TPR, y = PPV, color = threshold)) +
   scale_color_gradient(
     name = "Activity threshold (PIM)",
     limits= c(log10(5),log10(50)),
    guide = guide_colorbar(title.position = "top",
                           title.hjust = 0.5))  +
  ggnewscale::new_scale_color() + #need this to reset the color scale for the next plot
  #Adding the arrow to show directionality (for illuminance)
#  geom_segment(aes(x = 0.6,
 #                  y = 0.54,
 #                  xend = 0.63,
  #                 yend = 0.27),
   #            arrow = arrow(length=unit(.2, "cm")), lwd = 0.8) +
  #Second plot: light
  geom_point(data = prcurve_medi_sm, aes(x = TPR, y = PPV, colour = threshold)) +
   scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(log10(1),log10(10)),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  #Adding the arrow to show directionality (for activity)
 # geom_segment(aes(x = 0.49,
 #                  y = 0.165,
 #                  xend = 0.57,
 #                  yend = 0.115),
 #              arrow = arrow(length=unit(.2, "cm")), lwd = 0.8) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Precision-recall curve for detection of \nnon-wear intervals based on activity \nand illuminance thresholds") +
  #Adding the symbols for the arrows as text and the label
#  annotate("text",
 #          x = 0.64,
  #         y = 0.45,
   #        label = sprintf('\u2191')) + #arrow pointing upwards for Illuminance
  #geom_text(aes(x= 0.80,
   #             y = 0.445,
    #            label = "Illuminance")) + #label of illuminance
  #annotate("text",
   #        x = 0.28,
    #       y = 0.14,
     #      label = sprintf('\u2191')) + #arrow pointing upwards for activity
  #geom_text(aes(x = 0.39,
   #             y = 0.135,
    #            label = "Activity")) + #label for activity
  #geom_text(aes(x=0.215,
   #             y = 0.535,
    #            label = "Baseline classifier"),
     #       color = "darkgrey",
      #      size = 3.8) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 8),
        legend.key.size = unit(4, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal") +
  coord_fixed(ratio = 1) 

#Save the plot
ggsave(filename = "prc_smoothed_grouped.png",
       plot = prc_smoothed,
       height = 5.5, 
       width = 5,
       dpi = 600,
       bg = "white",
       path= "D:/cyepi/code/outputs/light_activity_prc")
```

##Now let's plot the smoothed prc for each individual
Note that you have to run the cluster detection another time, but grouping by Id and classification (see lines 118 and 203), before running this chunk. 
```{r}
prc_multiplot <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  #First plot: activity
  geom_point(data = prcurve_pim_sm, aes(x = TPR, y = PPV, color = threshold)) +
   scale_color_gradient(
     name = "Activity threshold (PIM)",
     low = "blue", high = "lightblue",
     limits= c(log10(5),log10(50)),
    guide = guide_colorbar(title.position = "top",
                           title.hjust = 0.5)) +
  facet_wrap(~ Id) +
  ggnewscale::new_scale_color() +
  #Second plot: light
  geom_point(data = prcurve_medi_sm, aes(x = TPR, y = PPV, colour = threshold)) +
   scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(log10(1),log10(10)),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  facet_wrap(~Id) +
  geom_text(data = per_bag_use, aes(x = 0.5, y = 0.95, label = paste("Bag use:",compliance_per,"%")), size = 4.2) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Precision-recall curve for detection of \nnon-wear intervals based on activity \nand illuminance thresholds") +
   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 16),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 14),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 8),
        legend.key.size = unit(4, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal") +
   coord_fixed(ratio = 1)

#Save the plot
ggsave(filename = "prc-multiplot-smoothed-labelled.png",
       plot = prc_multiplot,
       height = 10, 
       width = 10,
       dpi = 600,
       bg = "white",
       path= "D:/cyepi/code/outputs/light_activity_prc")

```

##Taking a step back: what if we maximised for each individual's max?
##Step 3: Maximise to max per dataset

```{r}
norm.rolled.df <- rolled.df %>%
  group_by(Id) %>%
  mutate(max_pim = max(rolled_median_PIM, na.rm = TRUE),
         max_medi = max(rolled_median_MEDI, na.rm = TRUE),
         norm_rolled_PIM = rolled_median_PIM/max_pim,
         norm_rolled_MEDI = rolled_median_MEDI/max_medi)
```

#Now apply the cluster detection to the normalised df 
```{r}
#Create a sequence of PIM thresholds at 5 unit steps
pimthresholds <- c(5,10, 15, 20, 25, 30, 35, 40, 45, 50)

log_thr_pim <- log10(pimthresholds)

#Empty list to store classification results
prc_norm_pim_list_sm <- list()
  
  for (threshold in log_thr_pim) {

  pim_norm_clust_sm <- norm.rolled.df %>%
    ungroup() %>%
    filter(!State == "sleep" ) %>%
    filter(!is.na(rolled_median_PIM)) %>% #account for the start and end of the rolling median, which has NAs   
   # Replace `low_activity` by the (logical) variable for which the clusters 
   # should be found.
   mutate(low_pim = norm_rolled_PIM < threshold) %>% 
   # Nest by groups, e.g., participant, period. Replace with grouping variables 
   # in your data, such that the function is applied separately per consecutive 
   # timeseries (e.g., from one subject during one period).
    nest_by(Id) %>% 
   mutate(
    # Replace `low_activity` by the (logical) variable for which the clusters 
    # should be found.
    data = list(
      data_find_clusters(
        data, 
        low_pim, 
        min_length = "10 mins", 
        max_interrupt = "1 mins", 
        cluster_name = "low_pim_cluster"
      )
    )
  ) %>%
  unnest(cols = data) %>%
  ungroup() 


  pimclusters_norm_clean_sm <- pim_norm_clust_sm %>%
   select(Id, Datetime, State, bag, is_low_pim_cluster) %>%
    mutate(
      State = case_when(
      State == "on" ~ 1, 
      State == "off" ~ 0),
    is_low_pim_cluster = case_when(
      is_low_pim_cluster == TRUE ~ 0,
      is_low_pim_cluster == FALSE ~ 1
    )) 

#We want to build a precision recall curve for our classifier algorithm, which, based o our "Ground truth", i.e. the wear log state values (on = 1 and off = 1), classifies the performance of our algorithm for detecting non-wear as follows:
#1) State = 0, cluster = 0 -> true positive
#2) State = 0, cluster = 1 -> false negative
#3) State = 1, cluster = 1 -> false positive
#4) State = 1, cluster = 0 -> true negative

  pimclusters_norm_clean_sm <- pimclusters_norm_clean_sm %>%
   mutate(classification = case_when(
      State == 0 & is_low_pim_cluster == 0 ~ "TP",
      State == 0 & is_low_pim_cluster == 1 ~ "FN",
      State == 1 & is_low_pim_cluster == 0 ~ "FP",
      State == 1 & is_low_pim_cluster == 1 ~ "TN",
     .default = NA_character_))

  prc_pim_norm_sm <- pimclusters_norm_clean_sm %>%
    group_by(classification) %>%
    summarise(count = n())  %>%
    pivot_wider(names_from = classification, values_from = count, values_fill = list(count=0)) %>%
  # Replace missing FN and TN with 0
    mutate(FN = replace_na(FN, 0),
         TN = replace_na(TN, 0),
         TP = replace_na(TP, 0),
         FP = replace_na(FP, 0)) %>%
    mutate(TPR = TP/(TP+FN), #this is the formula for TPR (true positive rate)
           FPR = FP/(FP+TN), #this is the formula for the FPR (false positive rate)
           PPV = TP/(TP+FP), #positive predictive value
           NPV = TN/(FN+TN), #negative predictive values
           threshold = threshold)  #adding manually which threshold I am considering here 

  #Add the result to the list
   prc_norm_pim_list_sm[[as.character(threshold)]] <- prc_pim_norm_sm
  }

#Turn the list with the results into a df 
prcurve_pim_norm_sm<- bind_rows(prc_norm_pim_list_sm)

#Turn thresholds from character to factor for plotting
prcurve_pim_norm_sm$threshold <- as.numeric(prcurve_pim_norm_sm$threshold)
```


```{r}
#Define threholds for cluser detection algorithm
medithresholds <- c(1,2,3,4,5,6,7,8,9,10)
log_thr_medi <-log10(medithresholds)

#Empty list to store classification results
prc_medi_list_sm <- list()

for (threshold in log_thr_medi) {

medi_clusters_sm <- norm.rolled.df %>%
  ungroup() %>%
  filter(!State == "sleep") %>%
  filter(!is.na(rolled_median_MEDI)) %>% #account for the start and end of the rolling median, which has NAs
  # Replace `low_activity` by the (logical) variable for which the clusters 
  # should be found.
  mutate(low_medi = norm_rolled_MEDI < threshold) %>% 
  # Nest by groups, e.g., participant, period. Replace with grouping variables 
  # in your data, such that the function is applied separately per consecutive 
  # timeseries (e.g., from one subject during one period).
  nest_by(Id) %>% 
  mutate(
    # Replace `low_activity` by the (logical) variable for which the clusters 
    # should be found.
    data = list(
      data_find_clusters(
        data, 
        low_medi, 
        min_length = "10 mins", 
        max_interrupt = "1 mins", 
        cluster_name = "low_medi_cluster"
      )
    )
  ) %>%
  unnest(cols = data) %>%
  ungroup() 


mediclusters_clean_sm <- medi_clusters_sm %>%
  select(Id, Datetime, State, is_low_medi_cluster) %>%
  mutate(
    State = case_when(
      State == "on" ~ 1, 
      State == "off" ~ 0),
    is_low_medi_cluster = case_when(
      is_low_medi_cluster == TRUE ~ 0,
      is_low_medi_cluster == FALSE ~ 1
    )) 

#We want to build a precision recall curve for our classifier algorithm, which, based o our "Ground truth", i.e. the wear log state values (on = 1 and off = 1), classifies the performance of our algorithm for detecting non-wear as follows:
#1) State = 0, cluster = 0 -> true positive
#2) State = 0, cluster = 1 -> false negative
#3) State = 1, cluster = 1 -> false positive
#4) State = 1, cluster = 0 -> true negative

mediclusters_clean_sm <- mediclusters_clean_sm %>%
  mutate(classification = case_when(
    State == 0 & is_low_medi_cluster == 0 ~ "TP",
    State == 0 & is_low_medi_cluster == 1 ~ "FN",
    State == 1 & is_low_medi_cluster == 0 ~ "FP",
    State == 1 & is_low_medi_cluster == 1 ~ "TN",
    .default = NA_character_))

prc_medi_sm <- mediclusters_clean_sm %>%
  group_by(classification) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = classification, values_from = count, values_fill = list(count=0)) %>%
  mutate(TPR = TP/(TP+FN), #true positive rate
         FPR = FP/(FP+TN), #false positive rate
         PPV = TP/(TP+FP), #positive predictive value
         NPV = TN/(FN+TN), #negative predictive value
         threshold = threshold)  #adding manually which threshold I am considering here 

 #Add the result to the list
  prc_medi_list_sm[[as.character(threshold)]] <- prc_medi_sm
}

#Turn the list with the results into a df 
prcurve_medi_sm<- bind_rows(prc_medi_list_sm)

#Turn thresholds from character to factor for plotting
prcurve_medi_sm$threshold <- as.numeric(prcurve_medi_sm$threshold)
```

## Where does the algorithm go wrong?
```{r}
#First, we need to create state changes from the mediclusters_clean data frame
prcurve_medi_sm <- prcurve_medi_sm %>% filter(threshold == log10(1))


medi_sc_sm <- prcurve_medi_sm %>%
 mutate(interval_group = consecutive_id(classification)) %>% #need to set default for first value 
  group_by(Id, interval_group) %>%
  summarise(classification = first(classification), #keep only first value for each Id and interval_group
            xmin = first(Datetime),
            xmax = last(Datetime)) %>%
  ungroup()


##List of all unique participant IDs
participant_ids <- unique(rolled.df$Id)

##Create a for loop to generate a plot for each participant 
for (selected_id in participant_ids) {
  
 subset_df <- rolled.df %>%
  filter(Id == selected_id)


##Calculate the limits of the shaded areas according to the Wear Log
alg.times <- medi_sc_sm %>%
  filter(Id == selected_id) %>%
  mutate(xmax = (xmax + lubridate::seconds(10)), #need to add 10 seconds to match next start time
         Day.data.min = xmin %>% format(format = "%a"), #lower limit of the shaded area, needed for geom_rect
         Day.data.max = xmax %>% format(format = "%a"), #upper limit of shaded area, needed for geom_rect
         xmin = xmin %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         xmax = xmax %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         Day.data.min = factor(Day.data.min, levels = unique(Day.data.min)),
         Day.data.max = factor(Day.data.max, levels = unique(Day.data.max)))

##Some intervals on the Wear log that we want to plot cross day boundaries. We thus need to split them in 2 intervals, since we plot each day on a different facet of the plot 

#Step 1: collect the part of the data that does not cross day boundaries
alg.times1 <-
  alg.times %>%
  filter(Day.data.min == Day.data.max)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
alg.times2 <-
alg.times %>% filter(Day.data.min != Day.data.max) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        xmin=c(xmin, hms::as_hms("00:00:00")),
        xmax=c(hms::as_hms("23:59:59"), xmax),
        Day.min=c(Day.data.min, Day.data.max),
        Day.max=c(Day.data.min, Day.data.max)
      )
    )
  ) %>%
  select(classification, new_entry) %>%
  unnest(new_entry)

#Step 3: Combine the two again in a single df
alg.combined <- 
  bind_rows(alg.times1 %>% rename(Day.min = Day.data.min),
            alg.times2) %>%
  select(-Day.max) %>%
  rename(Day.data = Day.min) 

alg.combined$classification <- factor(alg.combined$classification, levels = c("TP", "TN", "FP", "FN", "NA")) #turn State into factor

p <- gg_day(data = subset_df, #base plot defaults to plotting mEDI on y axis
            linewidth = 0.1, 
            aes_col = State, 
            geom = "line",
            group = date(Datetime),
            format.day = "%a",
            x.axis.label = "Time", 
            y.axis.label = "Illuminance (mEDI, lx)",
            title = paste("p", selected_id, "light levels and algorithm performance")) +
  geom_rect(data=alg.combined, aes(xmin= xmin, xmax = xmax, ymin =0, ymax = Inf, fill = classification), #adding classification output
                    inherit.aes = FALSE,
                    alpha = 0.15) +
    scale_fill_manual(
     name = "Algorithm classification",
     values = c("TP" = "darkgreen",
                "TN" = "#0072B2",
                "FP" = "darkred",
                "FN" = "darkgoldenrod2"),
                na.value = "white") +
  theme(plot.title = element_text(hjust = 0.5, size = 24, face = "plain"),
        axis.text.x = element_text(size = 16),
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14))

filename <- paste0("p", selected_id, "alg_perfom_sm.png")
ggsave(filename = filename,
       plot = p,
       width = 10,
       height = 14,
       dpi = 600,
       bg = "white",
       path= "G:/cyepi/code/outputs/light_activity_prc/smoothed/medi_alg_perf")
}
```


