---
title: "trblshoot_prc_medi"
output: html_document
date: "2024-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Fine tuning the input parameters for finding clusters of low illuminane
In this script, we produce precision-recall curves for each input parameter of the data_find_cluster function, in order to understand which parameters will work better to detect clusters of low illuminance. The data_find_cluster function takes three input parameters:
1. A threshold for the illuminance values (mEDI)
2. A minimum length of the interval 
3. A maximum interruption of the interval for values that are above threshold

In the chunks that follow, we vary each of these three parameters while keeping the other constant.

### First, import functions - if not in the environment yet
```{r}
# Fuction to identify clusters
base::source(here::here("03_analysis", "functions", "find_clusters_fun.R"))

#Function to generate aggregate pr curve
base::source(here::here("03_analysis", "functions", "generate_prc_funs.R"))
```


## Step 1: vary mEDI thresholds
Minimum interval length and maximum interruption remain constant.
```{r}

medithresholds <- seq(1, 10, by = 1)

#Empty list to store classification results
prc_m_list <- list()

for (threshold in medithresholds) {
  
  prc_result <- generate_prc(
    dataset = df.LL.nosleep,
    low_var = "MEDI",
    min_length = 60, #10 minutes (n of observations)
    max_interrupt = 0, #1 minute (n of observations)
    threshold = threshold
  )
  
  # Add the current threshold as a column in prc_m_result
  prc_result <- prc_result %>%
    mutate(threshold = threshold) # Add threshold column
  
  prc_m_list[[as.character(threshold)]] <- prc_result
  
}

#Turn list into df
prc_medi_1 <- bind_rows(prc_m_list)
#Turn threshold to numeric for plotting
prc_medi_1$threshold <- as.numeric(prc_medi_1$threshold)


#Calculate F1 score
prcmedi1_fscore <- prc_medi_1 %>%
  mutate(f1_score = (2 * PPV * TPR)/(PPV + TPR))

#Let's plot to see what the result looks like

prc_medi_thresholds <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  geom_point(data = prc_medi_1, aes(x = TPR, y = PPV, colour = threshold)) +
  scale_colour_gradient(
     name = "mEDI threshold (lx)",
     low = "red", high = "orange",
     limits= c(1,10),
     guide = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Illuminance threshold \nbased detection") +
   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.key.size = unit(8, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal",
        plot.margin = unit(c(0,0.5,0,0), unit ="cm")) +
   coord_fixed(ratio = 1)

ggsave(filename = "prc_medi_thresholds.png",
       plot = prc_medi_thresholds,
       height = 5.5, 
       width = 6,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "supplementary"))

```

## What does this tell us?
Compared to higher illuminance values of <10 mEDI lux, a lower illuminance value of <1 mEDI lux leads to higher precision (i.e. n of instances that were predicted as non-wear and are actually non-wear): 0.7237623 (1 lux) compared to 0.3404790 (10 lux). This increase in precision comes at a minimal cost for recall (i.e. n of instances correctly identified as non-wear out of all actual non-wear instances): 0.8111686 (1 lux) to 0.8918214 (10 lux). 


## Step 2: vary minimum length of interval 
Illuminance threshold and maximum intteruption remain constant.
```{r}
#We create min_length ranging from 2 to 11 minutes
min_lengths <- seq(6, 60, by = 6)
#From the previous plot, we know that lower mEDI thresholds are better so we keep it to 1
medi_threshold = 1
#We keep the max_interrupt at 0 minutes
max_interrupt = 0 # 0 minute


#Empty list to store classification results
prc_m_list <- list()

for (min_length in min_lengths) {
  
  prc_result <- generate_prc(
    dataset = df.LL.nosleep,
    low_var = "MEDI",
    min_length = min_length,
    max_interrupt = max_interrupt,
    threshold = medi_threshold
  )
  
  # Add the current min_length as a column in prc_m_result
  prc_result <- prc_result %>%
    mutate(min_length = min_length) # Add min_length column
  
  prc_m_list[[as.character(min_length)]] <- prc_result
}

#Turning this into a df 
prc_medi_2 <- bind_rows(prc_m_list)

#Turning min_length into a factor for plotting
prc_medi_2$min_length <- as.factor(prc_medi_2$min_length) 

#Calculate F1 score
prcmedi2_fscore <- prc_medi_2 %>%
  mutate(f1_score = (2 * PPV * TPR)/(PPV + TPR))

#Let's plot to see the results
prc_medi_length <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  geom_point(data = prc_medi_2, aes(x = TPR, y = PPV, colour = min_length)) +
  scale_colour_discrete(
     name = "Min. length (min)",
     labels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Minimum length \nbased detection",
       ) +
   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.key.size = unit(8, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal",
        plot.margin = unit(c(0,0.5,0,0), unit ="cm")) +
   coord_fixed(ratio = 1)

ggsave(filename = "prc_medi_length.png",
       plot = prc_medi_length,
       height = 5.5, 
       width = 6,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "supplementary"))

```

## What does this tell us?
Compared to low minimum length values (e.g. 1 minute), higher minimum length values of 10 minutes lead to  higher precision (i.e. n of instances that were predicted as non-wear and are actually non-wear): 0.7237623 (10 minutes) compared to 0.5940958 (1 minute). This increase in precision comes at a minor cost for recall: 0.8111686 (10 minutes) compared to 0.8241462 (1 minute). 

## Step 3: vary max interruption length 
Illuminance threshold and minimum length remain constant.
```{r}
#Create a max_interrupt ranging from 0 to 9
max_interrupts <- seq(0, 54, by = 6)
#From the previous plot, we know that lower mEDI thresholds are better so we keep it to 2
medi_threshold = 1
#We know also know that we shoudl keep the min_lenght to 10 minutes 
min_length = 60 # 10 minutes


#Empty list to store classification results
prc_m_list <- list()

for (max_int in max_interrupts) {
  
  prc_result <- generate_prc(
    dataset = df.LL.nosleep,
    low_var = "MEDI",
    min_length = min_length,
    max_interrupt = max_int,
    threshold = medi_threshold
  )
  
  # Add the current max_interrupt as a column in prc_result
  prc_result <- prc_result %>%
    mutate(max_int = max_int) # Add max_interrupt column
  
  prc_m_list[[as.character(max_int)]] <- prc_result
}

prc_medi_3 <- bind_rows(prc_m_list)

#Turning max_interrupt into a factor for plotting
prc_medi_3$max_int <- as.factor(prc_medi_3$max_int) 

#Calculate F1 score
prcmedi3_fscore <- prc_medi_3 %>%
  mutate(f1_score = (2 * PPV * TPR)/(PPV + TPR))

#Let's plot to see the results
prc_medi_interrupt <- ggplot() +
   xlim(0,1) + 
   ylim(0,1) +
  geom_abline(slope = 0, intercept = 0.5, linetype = "dashed", color = "darkgrey") + #add a flat line which represents a baseline classifier
  geom_point(data = prc_medi_3, aes(x = TPR, y = PPV, colour = max_int)) +
  scale_colour_discrete(
     name = "Max. interrupt (min)",
     labels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9)) +
  labs(x="True positive rate (recall)",
       y = "Positive predictive value \n(precision)",
       title = "Maximum interruption \nbased detection") +
   theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 20),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.key.size = unit(8, "mm"),
        legend.position = "bottom",
        legend.box = "horizontal",
        plot.margin = unit(c(0,0.5,0,0), unit ="cm")) +
   coord_fixed(ratio = 1)

ggsave(filename = "prc_medi_interrupt.png",
       plot = prc_medi_interrupt,
       height = 5.5, 
       width = 6.3,
       dpi = 600,
       bg = "white",
       path= here::here("outputs", "supplementary"))

```

## What does this tell us?
Different interruption lengths of the non-wear interval do not affect the precision or the recall of the cluster detection function. 

## One plot with all the three curves 
```{r}

library(patchwork)

multiplot <- cowplot::plot_grid(prc_medi_length, prc_medi_interrupt, prc_medi_thresholds,
                                labels = c("A", "B", "C"),
                                ncol = 3,
                                nrow = 1,
                                align = "hv")

ggsave(filename = "medi_prc_parameters_multiplot_1.png",
       plot = multiplot,
       width = 14, 
       height = 6,
       dpi = 600, 
       bg = "white",
       path = here::here("outputs", "supplementary"))

```
## Conclusion
The values to be used are:
1. Illuminance threshold: 1 lux
2. Minimum length: 10 minutes 
3. Maximum interruption: 0 minutes 
