---
title: "nw_visual_insp"
author: "Carolina Guidolin"
date: "2024-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Visual inspection of the data and different sources of non-wear
As part of the non-wear analysis of this light logger dataset, one of the first steps to perform is just to visually inspect the data. This gives us an insight in what analyses would be possible across the dataset, and inspires research questions that we had perhaps not considered. The following code should be run after all scripts in the folder 01_import and 02_datapreparation. We want to plot the following:
- Activity (PIM)
- Light (mEDI)
- Button presses participants performed any time they removed or placed the light glasses on 
- Information on the Wear log (light glasses on/off/off during sleep)

##Creating a plot that represents the 4 variables mentioned above for each day of the week, for each participant
Note that since both mEDI and PIM are continuous measures and the raw data has a 10s resolution, plotting them in the same graph would lead to a lot of noise. Hence, we create two plots for each participant
1) pX_activity: this plot plots PIM values, Wear log intervals, button presses
2) pX_light: this plot plots mEDI values, Wear log intervals, button presses

Since we have all of this information already loaded after we integrated the sources of non-wear in the df dataset.LL.wrlg, we will use this instead of leading each each light logger dataset again from each participant.

## Creating plot 1) pX_activity
```{r}

##List of all unique participant IDs
participant_ids <- unique(dataset.LL.wrlg$Id)

selected_id <- "201"

##Create a for loop to generate a plot for each participant 
for (selected_id in participant_ids) {

subset_df <- dataset.LL.wrlg %>%
  filter(Id == selected_id) %>%
  select(Id, Datetime, EVENT, PIM, MEDI, State)
  

##Create new column for the y-values (PIM) of points where EVENT==1
subset_df$eventpoint <- ifelse(subset_df$EVENT == 1, subset_df$PIM, NA)

##Calculate the limits of the shaded areas according to the Wear Log
state.times <- wrlg_int %>%
  filter(Id == selected_id & !is.na(State)) %>%
  filter(row_number() != n()) %>%
  mutate(xmin = int_start(Interval),
         xmax = int_end(Interval),
         Day.data.min = xmin %>% format(format = "%a"), #lower limit of the shaded area, needed for geom_rect
         Day.data.max = xmax %>% format(format = "%a"), #upper limit of shaded area, needed for geom_rect
         xmin = xmin %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         xmax = xmax %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         Day.data.min = factor(Day.data.min, levels = unique(Day.data.min)),
         Day.data.max = factor(Day.data.max, levels = unique(Day.data.max)))

##Some intervals on the Wear log that we want to plot cross day boundaries. We thus need to split them in 2 intervals, since we plot each day on a different facet of the plot 

#Step 1: collect the part of the data that does not cross day boundaries
state.times1 <-
  state.times %>%
  filter(Day.data.min == Day.data.max)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
state.times2 <-
state.times %>% filter(Day.data.min != Day.data.max) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        xmin=c(xmin, hms::as_hms("00:00:00")),
        xmax=c(hms::as_hms("23:59:59"), xmax),
        Day.min=c(Day.data.min, Day.data.max),
        Day.max=c(Day.data.min, Day.data.max)
      )
    )
  ) %>%
  select(Id, State, new_entry) %>%
  unnest(new_entry)

#Step 3: Combine the two again in a single df
states.combined <- 
  bind_rows(state.times1 %>% rename(Day.min = Day.data.min),
          state.times2) %>%
  select(-Day.max) %>%
  rename(Day.data = Day.min) 


#Create the plot wit activity (PIM), button presses (eventpoint) and wearlog intervals (State)

states.combined$State <- factor(states.combined$State) #turn State into factor

p <- gg_day(data = subset_df,
            y.axis = PIM, #base plot containing PIM
            linewidth = 0.05, 
            color = "black", 
            group = date(Datetime),
            geom = "line", 
            y.scale = "identity",
            y.axis.label = "Activity (PIM)",
            y.axis.breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000),
            format.day = "%a",
            x.axis.label = "Time", 
            title = paste("p", selected_id, "activity and non-wear time")) +
  geom_point(aes(y = eventpoint), #adding button press as points
             color = "red", 
             size = 0.8) +
  geom_rect(data=states.combined, aes(xmin= xmin, xmax = xmax, ymin =0, ymax = Inf, fill = State), #adding wear log
                    inherit.aes = FALSE,
                    alpha = 0.25) +
   scale_fill_manual(
     name = "State",
     values = c("off" = "#16439C", "on" = "palegreen4", "sleep" = "darkgoldenrod2"),
     labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  
  theme(plot.title = element_text(hjust = 0.5)) 

filename <- paste0("p", selected_id, "_activity.png")
ggsave(filename = filename,
       plot = p,
       width = 10,
       height = 14,
       dpi = 600,
       bg = "white",
       path= "G:/cyepi/code/outputs/visual_insp/activity")
}


```

###Creating plot 2) pX_light
Note that this code is the same as above, but is used slightly adjusted to plot light (mEDI) instead of activity
```{r}
##List of all unique participant IDs
participant_ids <- unique(dataset.LL.wrlg$Id)

##Create a for loop to generate a plot for each participant 
for (selected_id in participant_ids) {
subset_df <- dataset.LL.wrlg %>%
  filter(Id == selected_id) %>%
  select(Id, Datetime, EVENT, PIM, MEDI, State)

##Create new column for the y-values (medi) of points where EVENT==1
subset_df$eventpoint <- ifelse(subset_df$EVENT == 1, subset_df$MEDI, NA)

##Calculate the limits of the shaded areas according to the Wear Log
state.times <- wrlg_int %>%
  filter(Id == selected_id & !is.na(State)) %>%
  filter(row_number() != n()) %>%
  mutate(xmin = int_start(Interval),
         xmax = int_end(Interval),
         Day.data.min = xmin %>% format(format = "%a"), #lower limit of the shaded area, needed for geom_rect
         Day.data.max = xmax %>% format(format = "%a"), #upper limit of shaded area, needed for geom_rect
         xmin = xmin %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         xmax = xmax %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         Day.data.min = factor(Day.data.min, levels = unique(Day.data.min)),
         Day.data.max = factor(Day.data.max, levels = unique(Day.data.max)))

##Some intervals on the Wear log that we want to plot cross day boundaries. We thus need to split them in 2 intervals, since we plot each day on a different facet of the plot 

#Step 1: collect the part of the data that does not cross day boundaries
state.times1 <-
  state.times %>%
  filter(Day.data.min == Day.data.max)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
state.times2 <-
state.times %>% filter(Day.data.min != Day.data.max) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        xmin=c(xmin, hms::as_hms("00:00:00")),
        xmax=c(hms::as_hms("23:59:59"), xmax),
        Day.min=c(Day.data.min, Day.data.max),
        Day.max=c(Day.data.min, Day.data.max)
      )
    )
  ) %>%
  select(Id, State, new_entry) %>%
  unnest(new_entry)

#Step 3: Combine the two again in a single df
states.combined <- 
  bind_rows(state.times1 %>% rename(Day.min = Day.data.min),
          state.times2) %>%
  select(-Day.max) %>%
  rename(Day.data = Day.min) 


#Create the plot with light (medi)), button presses (eventpoint) and wearlog intervals (State)

states.combined$State <- factor(states.combined$State) #turn State into factor

p <- gg_day(data = subset_df, #base plot defaults to plotting mEDI on y axis
            linewidth = 0.05, 
            color = "black", 
            geom = "line",
            format.day = "%a",
            x.axis.label = "Time", 
            y.axis.label = "Illuminance (mEDI, lx)",
            title = paste("p", selected_id, "light exposure and non-wear time")) +
  geom_point(aes(y = eventpoint), #adding button press as points
             color = "red", 
             size = 0.8) +
  geom_rect(data=states.combined, aes(xmin= xmin, xmax = xmax, ymin =0, ymax = Inf, fill = State), #adding wear log
                    inherit.aes = FALSE,
                    alpha = 0.15) +
   scale_fill_manual(
     name = "State",
     values = c("on" = "#0072B2", "off" = "darkred", "sleep" = "darkgoldenrod2"),
     labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  theme(plot.title = element_text(hjust = 0.5)) 

filename <- paste0("p", selected_id, "_light.png")
ggsave(filename = filename,
       plot = p,
       width = 10,
       height = 14,
       dpi = 600,
       bg = "white",
       path= "D:/cyepi/code/outputs/visual_insp/light")
}

```

## Only visualising this for one participant
For Figure 4 of the paper, we would like to create a single plot for one representative participant. We choose participant 201 for this purpose. 
```{r}

selected_id <- "201"

subset_df <- dataset.LL.wrlg %>%
  filter(Id == selected_id) %>%
  select(Id, Datetime, EVENT, PIM, MEDI, State)

##Create new column for the y-values (PIM) of points where EVENT==1
subset_df$eventpoint <- ifelse(subset_df$EVENT == 1, subset_df$MEDI, NA)

##Calculate the limits of the shaded areas according to the Wear Log
state.times <- wrlg_int %>%
  filter(Id == selected_id & !is.na(State)) %>%
  filter(row_number() != n()) %>%
  mutate(xmin = int_start(Interval),
         xmax = int_end(Interval),
         Day.data.min = xmin %>% format(format = "%a"), #lower limit of the shaded area, needed for geom_rect
         Day.data.max = xmax %>% format(format = "%a"), #upper limit of shaded area, needed for geom_rect
         xmin = xmin %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         xmax = xmax %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         Day.data.min = factor(Day.data.min, levels = unique(Day.data.min)),
         Day.data.max = factor(Day.data.max, levels = unique(Day.data.max)))

##Some intervals on the Wear log that we want to plot cross day boundaries. We thus need to split them in 2 intervals, since we plot each day on a different facet of the plot 

#Step 1: collect the part of the data that does not cross day boundaries
state.times1 <-
  state.times %>%
  filter(Day.data.min == Day.data.max)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
state.times2 <-
state.times %>% filter(Day.data.min != Day.data.max) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        xmin=c(xmin, hms::as_hms("00:00:00")),
        xmax=c(hms::as_hms("23:59:59"), xmax),
        Day.min=c(Day.data.min, Day.data.max),
        Day.max=c(Day.data.min, Day.data.max)
      )
    )
  ) %>%
  select(Id, State, new_entry) %>%
  unnest(new_entry)

#Step 3: Combine the two again in a single df
states.combined <- 
  bind_rows(state.times1 %>% rename(Day.min = Day.data.min),
          state.times2) %>%
  select(-Day.max) %>%
  rename(Day.data = Day.min)


#Create the plot wit activity (PIM), button presses (eventpoint) and wearlog intervals (State)

states.combined$State <- factor(states.combined$State) #turn State into factor

p <- gg_day(data = subset_df, #base plot defaults to plotting mEDI on y axis
            linewidth = 0.03, 
            color = "black", 
            geom = "line",
            #alpha = 0.7,
            group = date(Datetime),
            format.day = "%a",
            x.axis.label = "Local time (HH:MM)", 
            y.axis.label = "Illuminance (mEDI, lx)",
            title = paste("Light exposure and wear status")) +
  geom_point(aes(y = eventpoint, color = "Button press"), #adding button press as points
             size = 0.8) +
  geom_rect(data=states.combined, aes(xmin= xmin, xmax = xmax, ymin =0, ymax = Inf, fill = State), #adding wear log
                    inherit.aes = FALSE,
                    alpha = 0.15) +
   scale_fill_manual(
     name = "Wear log state",
     values = c("on" = "#0072B2", "off" = "darkred", "sleep" = "darkgoldenrod2"),
     labels = c("Wake 'off'", "Wake 'on'", "Sleep 'off'")) +
  scale_color_manual(
    name = "Event",
    values = c("Button press" = "red"),
    labels = c("Button press")) +
  theme(plot.title = element_text(size = 24, face = "plain"),
        axis.text.x = element_text(size = 16),
        axis.title = element_text(size = 18),
        axis.text.y = element_text(size = 16),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.position = "inside", 
        legend.position.inside = c(0.08, 0.935),
        legend.box.background = element_rect(fill = "white", colour = "darkgrey"),
        legend.margin = margin(c(1,1,1,1)))

ggsave(filename = "results_fig01.png",
       plot = p,
       width = 11,
       height = 14,
       dpi = 600,
       bg = "white",
       path= "H:/nonwear_detection/preprint_figures/results/fig01")

```

