---
title: "wrist_eye_corrs"
author: "Carolina Guidolin"
date: "2024-07-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##NOTE: to be run after import_ll and import_activity 
```{r}
dataset.LL.all.1 <- dataset.LL.all %>%
  rename(timestamp = Datetime) #rename Datetime column so we do not get confused between the two files

data.table::setDT(dataset.LL.all.1)[, join_date := timestamp] #convert to data.table and append original column that we want to keep
data.table::setDT(dataset.acttrust.clean)[, join_date := Datetime] #convert to data.table and append original column that we want to keep


#For each ActTrust file, find the nearest (in time) ActLumus Datetime
acttrust_lumus_join <- dataset.LL.all.1[dataset.acttrust.clean, on = .(Id, join_date), roll = "nearest"] %>% #Perform a left join of the two datasets using data.table
  rename(actlumus_PIM = PIM, acttrust_PIM = i.PIM, actlumus_MEDI = MEDI, acttrust_light = i.LIGHT) 

#Because we will plot in log10, we need to eliminate any 0 values from both PIM and light
acttrust_lumus_join$actlumus_MEDI[acttrust_lumus_join$actlumus_MEDI == 0] <- 0.00001
acttrust_lumus_join$actlumus_PIM[acttrust_lumus_join$actlumus_PIM == 0] <- 0.00001
acttrust_lumus_join$acttrust_PIM[acttrust_lumus_join$acttrust_PIM == 0] <- 0.00001
acttrust_lumus_join$acttrust_light[acttrust_lumus_join$acttrust_light == 0] <- 0.00001
  
#See the concordance between wrist and eye level activity for each Id using a for loop

part_ids <- unique(acttrust_lumus_join$Id)

for (id in part_ids) {
  
  #Filter data for current id
  indiv_dataset <- subset(acttrust_lumus_join, Id == id)
  
  p <- ggplot(indiv_dataset, aes(x = actlumus_PIM, y = acttrust_PIM)) +
    geom_point() +
    labs(x = "Eye-level activity (PIM)", y = "Wrist-level activity (PIM)", title = paste0("PID_", id, "_activity")) +
    stat_cor(method = "pearson") +
    scale_y_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    scale_x_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    coord_fixed(ratio = 1) +
    theme_bw()
  
  print(p)
}
  
```

#See the concordance between wrist and eye level light for each Id using a for loop 
```{r}


part_ids <- unique(acttrust_lumus_join$Id)

for (id in part_ids) {
  
  #Filter data for current id
  indiv_dataset <- subset(acttrust_lumus_join, Id == id)
  
  p <- ggplot(indiv_dataset, aes(x = actlumus_MEDI, y = acttrust_light)) +
    geom_point(size = 1, colour = "darkblue", alpha = 0.7) +
    labs(x = "Eye-level light (mEDI)", y = "Wrist-level light", title = paste0("PID_", id, "_light")) +
    stat_cor(method = "pearson") +
    scale_y_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    scale_x_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    coord_fixed(ratio = 1) +
    theme_bw()
  
  print(p)
}

```

#Now, create a multi-panel lot where all of the individual plots are shown 
1) Illuminance
```{r}
light_corr <-
ggplot(acttrust_lumus_join, aes(x = actlumus_MEDI, y = acttrust_light)) +
    geom_point(size = 0.7, colour = "#fad47c", alpha = 0.7) +
    labs(x = "Eye-level illuminance (mEDI) [lx]", y = "Wrist-level illuminance [lx]", title = "Relationship between wrist- and eye-level illuminance") +
    stat_cor(method = "pearson", aes(label = after_stat(r.label))) +
    scale_y_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    scale_x_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
  theme(plot.title = element_text(hjust = 0.5)) +
    coord_fixed(ratio = 1) +
    theme_bw() +
  facet_wrap(~Id,
             ncol = 6)

ggsave("light_corr.pdf",
       plot = light_corr,
       height = 8,
       width = 8,
       dpi = 600,
       path = "G:/cyepi/code/outputs/visualinspection/wrist_eye_corrs")

```

2) Do the same for activity
```{r}
activity_corr <-
ggplot(acttrust_lumus_join, aes(x = actlumus_PIM, y = acttrust_PIM)) +
    geom_point(size = 0.7, colour = "#2A5B84", alpha = 0.7) +
    labs(x = "Eye-level activity [PIM]", y = "Wrist-level activity [PIM]", title = "Relationship between wrist- and eye-level activity") +
    stat_cor(method = "pearson", aes(label = after_stat(r.label))) +
    scale_y_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    scale_x_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
  theme(plot.title = element_text(hjust = 0.5)) +
    coord_fixed(ratio = 1) +
    theme_bw() +
  facet_wrap(~Id,
             ncol = 6)

ggsave("activity_corr.pdf",
       plot = activity_corr,
       height = 8,
       width = 8,
       dpi = 600,
       path = "G:/cyepi/code/outputs/visualinspection/wrist_eye_corrs")
```

## Once we have run the correlations, we want to visualise the distrubution of light exposure over the distribution of activity, for each participant.
We can do this using a qq plot. For each participant, you calculate the quantiles: 1%, 5%, 10% and so on. These give you a PIM or mEDI value below which 1%, 5% or 10% of the data is. You then plot the obtained values on two separate axes for each participant.
```{r}
#Creating an empty dataframe for later
all_quantiles_df <- data.frame()

#Creating a sequence for the probabilities we are interested in
probabilities <- seq(0.1, 1, length.out = 10) 

for (id in part_ids) {
  
  #Subsetting for that ID
  indiv_dataset <- subset(acttrust_lumus_join, Id == id)
  
  #Calculate the quantiles of mEDI and PIM
  quantiles_light <- quantile(indiv_dataset$actlumus_MEDI, probs = probabilities)
  quantiles_PIM <- quantile(indiv_dataset$acttrust_PIM, probs = probabilities)

  #Place this into a df
  quantiles_df <- data.frame(light = quantiles_light,
                             activity = quantiles_PIM,
                             quantile_index = seq_along(probabilities)*10)
  
  #Add a column for PID
  quantiles_df$id <- id
  
  #Print
  print(quantiles_df)
  
  #Append to bigger df 
 all_quantiles_df <- rbind(all_quantiles_df, quantiles_df)
 
 # Reset row numbers in the final dataframe to avoid having the index as percentages
  rownames(all_quantiles_df) <- NULL
  
}

```

Now that we have the df, we create a scatterplot of mEDI and PIM for each participant and visualise this as a multi plot with ~facet_wrap
```{r}
library(ggrepel)

quantiles_plot <- 
  ggplot(all_quantiles_df, aes(x=light, y=activity)) +
  geom_jitter(size=1.5) +
  labs(x="Eye-level illuminance (mEDI)", y="Wrist-level activity (PIM)", title = "Distribution of light and activity across participants") +
  scale_y_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
    scale_x_continuous(trans = "symlog",
                       breaks = c(-10^(5:0), 0, 10^(0:5)),
                       labels = function(x) format(x, scientific = FALSE, big.mark = " "),
                       limits = c(0, 10^5)) +
  geom_text_repel(aes(label = ifelse(
    quantile_index == "10" | quantile_index == "50" | quantile_index == "100", quantile_index,"")), #only added some of the labels, as adding all would make the plot too messy
            size= 3,
            hjust=-0.5,
            vjust = 0.5) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) +
  coord_fixed(ratio=1) +
  facet_wrap(~id,
             ncol=6)

ggsave("quantiles_plot.pdf",
       plot = quantiles_plot,
       height = 10,
       width = 10,
       dpi = 600,
       path = "G:/cyepi/code/outputs/visualinspection/wrist_eye_corrs")


```

