---
title: "plots_wrlg"
author: "Carolina Guidolin"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wear log plots

This script creates:
1) Plots to visualise how much time participants spent in each Wear log state across the week
2) Descriptive statistics on how much time participants spent in each Wear log state across the week
3) Plots to visualise when, across 24 hours, non-wear events (according to the wear log) occur
4) Plots to visualise non-wear time at weekend vs. weekdays


## 1) Plots to visualise how much time participants spend in each wear log state across the week.
We would like to visualise this as a distribution of time spent in each State, where the max x axis is 7 days (168 hours). Since 168 hours is not very indicative for someone looking at the graph, we also want to express this as a percentage of time, where 168 hours is 100%.

#Load necessary packes
```{r}
library(ggdist) # to create raincloud plots
library(patchwork) 
```


```{r}
#First, while all participants took part in the experiment for 7 days, they started at different times, and so it is important not just to approximate to 7 days (168 hours), but to calculate the specific 100% of "participation time" for each participant. For this, we first need to calculate the n of hours that each participant spends in the experiment.

exp_duration <- filtered_time %>%
  mutate(id_int = lubridate::interval(start, end),
         id_duration = lubridate::int_length(id_int) %>% as_hms()) %>% #tot hours of participation for each participant
  select(Id, id_duration)


#Time in each interval can be calculated from wrlg_int 
int_duration <- wrlg_int %>%
  group_by(Id) %>%
  #filter out the last "on" interval, since it goes to midnight of the final exp day, which is not actual data
  filter(row_number() != n()) %>% #this filters out the rows where row n is equal to tot number of rows, i.e. last entry
  ungroup() %>%
  #calculate length of each individual interval
  mutate(int_length = int_length(Interval) %>% as_hms()) %>% 
  #group by Id and State to then  summarise the total (i.e. across the whole week) duration of each state (on/off/sleep) 
  group_by(Id, State, .add = TRUE) %>%
  summarize(tot_intlength = sum(int_length) %>% as_hms()) %>%
  ungroup() %>%
  #The first states for each Id (midnight of previous day to study start) are NA, we want to filter them out since they are not active participation
  filter(!is.na(State))

#Normalise it to the total time participants spent in the experiment, expressed in percentage
int_duration <- int_duration %>%
  left_join(exp_duration, by ="Id") %>% #append the column containing total participation hours from the exp_duration df
  mutate(id_duration_n= as.numeric(tot_intlength)/as.numeric(id_duration), #convert to numeric as division betwen difftime objects is not supported, n stands for "normalised"
         id_duration_per = id_duration_n*100) #calculate percentage of time

#Turn State into a factor
as.factor(int_duration$State)

#Plot
int_duration$State <- factor(int_duration$State, levels = c("sleep", "off", "on"))

states_dist_jitter2 <- ggplot(data = int_duration, aes(x=id_duration_per, y = State, fill = State)) +
  scale_x_continuous(limits = c(0,100), expand = c(0.05,0.05)) +
  ggdist::stat_halfeye(
    aes(colour = State,
        fill = State),
    adjust = .5,
    justification = -.25,
    point_color = NA,
    interval_color = NA
    ) +
  geom_boxplot(
    aes(color = State),
    width = .3,
    alpha = .2,
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(colour = State),
    fill = "white",
    height = .1,
    alpha = .3,
    size = 1.2
  ) +
  scale_fill_manual(
    values = c("on" = "#0072B2", "off" = "darkred", "sleep" = "darkgoldenrod2")) +
  scale_color_manual(
    values = c("on" = "#0072B2", "off" = "darkred", "sleep" = "darkgoldenrod2")) +
  scale_y_discrete(labels = c("on" = "Wake 'on'", "off" = "Wake 'off'", "sleep" = "Sleep 'off'")) +
  theme_ggdist() +
  ggpubr::rremove("ylab") +
  ggpubr::rremove("y.ticks") +
 # ggpubr::rremove("xlab")+ 
  labs(title = "Wear status distribution across the week", x = "Percentage of time (%)") +
  coord_fixed(ratio = 14) + # reduce white space between categorical variables on y axis
  theme(legend.position = "none",
        plot.title = element_text(size = 18),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 16),
        axis.text.y = element_text(size = 14, colour = "black"))

#We want to add a 2nd axis to the plot, where hours, instead of % of time, are specified

int_duration$tot_intlength_hours <- as.numeric(int_duration$tot_intlength) / 3600
  
second_x_axis <- ggplot(data = int_duration, aes(x = tot_intlength_hours, y = 0.01)) +
  geom_blank() +  # Create an empty plot
  scale_x_continuous(limits = c(0, 166), breaks = c(0, 40, 80, 120, 166), 
                     expand = c(0.05,0.05)) +
  coord_fixed(ratio = 4) +
  theme_ggdist() +
  labs(x = "Time (hours)")  +
  theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  ggpubr::rremove("y.axis") +
  ggpubr::rremove("ylab") + 
  ggpubr::rremove("y.ticks") +
  ggpubr::rremove("y.text") 
  

# Combine the plots

state_dist_2axis_2 <- states_dist_jitter2/second_x_axis

state_dist_2axis_2 <- state_dist_2axis_2 + theme(plot.margin=margin(t = 0,
                                                                  b = 10,
                                                                  r = 10,
                                                                  l = 10))


```

## Save combined plot above
```{r}

ggsave(filename = "results_fig02a.svg",
       plot = state_dist_2axis_2,
       width = 7,
       height = 4,
       dpi = 600,
       path= "H:/nonwear_detection/preprint_figures/results/fig02")
```

## 2) Calculate descriptive statistics on how much time participants spent in each Wear log state across the week 
```{r}
#Calculating median time in each state

wrlg_state_summary <- int_duration %>%
  group_by(State) %>%
  summarise(mean = mean(id_duration_per),
            sd = sd(id_duration_per),
            min = min(id_duration_per),
            max = max(id_duration_per),
            median = median(id_duration_per),
            range = max-min) %>%
  ungroup()

```

## 3) We now want to visualise when, during the day, participants took the light logger on and/or off
```{r}
timing_off <- wrlg_int %>%
  filter(State == "off") %>%
  mutate(start = int_start(Interval),
         end = int_end(Interval),
         centre_point = as.POSIXct(start + difftime(end,start)/2, tz = tz),
         centre_point_h = hour(centre_point))

summary_timing <- timing_off %>%
  group_by(centre_point_h) %>%
  summarise(count = n())

timing_dens_plot <- ggplot(timing_off, aes(x = as.numeric(centre_point_h), y = after_stat(count))) +
  geom_histogram(color = "black", fill = "darkred", alpha = 0.6, position = "identity", bins = 24) +
  geom_density(color = "black", fill= "darkred", alpha = 0.3, outline.type = "upper") +
  labs(x = "Local time (HH:MM)", y = "Frequency of \nWake 'off' events", title = "Timing of non-wear episodes across the day") +
  scale_x_continuous(
    breaks = c(0, 2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24),
    labels = c("00:00", "02:00", "04:00", "06:00", "08:00", "10:00", "12:00", "14:00", "16:00", "18:00", "20:00", "22:00", "24:00"),
    expand = c(0.0,0.0)) +
  theme_ggdist() +
   theme(legend.position = "none",
        plot.title = element_text(size = 18),
        axis.text.x = element_text(size = 14, angle = 45, hjust = 1),
        axis.title = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        plot.margin=margin(t = 0,
                           b = 10,
                           r = 10,
                           l = 10)) +
  coord_fixed(ratio = 0.6) 

# Saving the plot
ggsave(filename = "results_fig02c.png",
       plot = timing_dens_plot,
       width = 7,
       height = 6,
       dpi = 600,
       path= "H:/nonwear_detection/preprint_figures/results/fig02")

```

## 4) Looking at weekday vs weekend difference
```{r}
int_wknd_wkday <- wrlg_int %>%
  group_by(Id) %>%
  #filter out the last "on" interval, since it goes to midnight of the final exp day, which is not actual data
  filter(State == "off") %>%
  ungroup() %>%
  #calculate length of each individual interval
  mutate(int_length = int_length(Interval) %>% as_hms(),
         int_start = int_start(Interval),
         int_end = int_end(Interval),
         day.start = int_start %>% format(format = "%a"), #format as day
         day.end = int_end %>% format(format = "%a"), #format as day
         int_start = int_start %>% hms::as_hms(), #transform xmin to time (currently Datetime)
         int_end = int_end %>% hms::as_hms(), #transform xmax to time (currently Datetime)
         day.start = factor(day.start, levels = unique(day.start)),
         day.end = factor(day.end, levels = unique(day.end)))

#Step 1: collect the part of the data that does not cross day boundaries
off_times1 <-
  int_wknd_wkday %>%
  filter(day.start == day.end)

#Step 2: Collect part of the data that does cross bounderies and split it by day, i.e. at midnight. 
off_times2 <-
int_wknd_wkday %>% filter(day.start != day.end) %>%
  rowwise() %>%
  mutate(
    new_entry = list(
      tibble(
        int_start=c(int_start, hms::as_hms("00:00:00")),
        int_end =c(hms::as_hms("23:59:59"), int_end),
        day.start.1=c(day.start, day.end),
        day.end.1=c(day.start, day.end)
      )
    )
  ) %>%
  select(Id, State, Interval, new_entry) %>%
  unnest(new_entry)

wknd_wkday.combined <- 
  bind_rows(off_times1 %>% rename(day.start.1 = day.start),
          off_times2) %>%
  select(-c(day.end.1, day.end, int_length)) %>%
  rename(Day.data = day.start.1) %>%
  mutate(off_length = as_hms(int_end - int_start),
         day_type = ifelse(Day.data %in% c("Sat", "Sun"), "weekend", "weekday")) %>%
  #group by Id and State to then  summarise the total (i.e. across the whole week) duration of each state (on/off/sleep) 
  group_by(Id, day_type) %>%
  summarize(off_length = sum(off_length) %>% as_hms()) %>%
  ungroup() %>%
  complete(Id, day_type, fill = list(off_length = as_hms(0)))

wknd_wkday <- wknd_wkday.combined %>%
  left_join(exp_duration, by ="Id") %>% #append the column containing total participation hours from the exp_duration df
  mutate(wknd_duration = hms::as_hms(48 * 3600),
         wk_duration = hms::as_hms(id_duration - wknd_duration),
         normalised_length = if_else(day_type == "weekend", as.numeric(off_length)/as.numeric(wknd_duration), 
                                     as.numeric(off_length)/as.numeric(wk_duration)),
         perc_length = normalised_length * 100)

# Turning day_type into factor for plotting
wknd_wkday$day_type <- factor(wknd_wkday$day_type, levels = c("weekend", "weekday"))

# Raincloud plot
 wknd_wk_dist <- ggplot(data = wknd_wkday, aes(x=perc_length, y = day_type, fill = day_type)) +
  scale_x_continuous(limits = c(0,100), expand = c(0.05,0.05)) +
  ggdist::stat_halfeye(
    aes(colour = day_type,
        fill = day_type),
    adjust = .5,
    justification = -.45,
    point_color = NA,
    interval_color = NA,
    scale = 0.5,
    height = 1
    ) +
  geom_boxplot(
    aes(color = day_type),
    width = 0.25,
    alpha = .2,
    outlier.shape = NA
  ) +
  geom_jitter(
    aes(colour = day_type),
    fill = "white",
    height = .1,
    alpha = .3,
    size = 1.2,
    na.rm = TRUE
  ) +
  scale_fill_manual(
    values = c("weekday" = "#bf0000", "weekend" = "#800000")) +
  scale_color_manual(
    values = c("weekday" = "#bf0000", "weekend" = "#800000")) +
  scale_y_discrete(labels = c("weekday" = "Weekday", "weekend" = "Weekend day")) +
  theme_ggdist() +
  ggpubr::rremove("ylab") +
  ggpubr::rremove("y.ticks") +
 # ggpubr::rremove("xlab")+ 
  labs(title = "Wake 'off' status distribution", x = "Percentage of time (%)") +
  coord_fixed(ratio = 18) + # reduce white space between categorical variables on y axis
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, size = 18),
        axis.text.x = element_text(size = 14),
        axis.title = element_text(size = 16),
        axis.text.y = element_text(size = 14, colour = "black"),
        plot.margin=margin(t = 0,
                           b = 10,
                           r = 10,
                           l = 10))
 

```

## Saving the plot
```{r}
 #Saving the plot
 ggsave(filename = "results_fig02b.png",
       plot =  wknd_wk_dist,
       width = 7,
       height = 4,
       dpi = 600,
       path= "H:/nonwear_detection/preprint_figures/results/fig02")

```

### Creating a multiplot with the three plots above - they will all go in figure 02
```{r}
library(patchwork)

fig02_stacked <- state_dist_2axis_2/wknd_wk_dist/timing_dens_plot 


# Adding annotations
fig02_multiplot <- fig02_stacked + patchwork::plot_annotation(
  tag_levels = list(c("A", "", "B", "C")) # adding empty string because patchwork would otherwise recognise the 2nd axis of plot A as an additional plot
) & theme(plot.tag = element_text(face = "bold")) + #need to add this with & operator to get labels in bold
  theme(plot.margin=margin(t = 0,
                           b = 0,
                           r = 0,
                           l = 0)) 


ggsave(filename = "results_fig02_all.png",
       plot = fig02_multiplot,
       width = 8,
       height = 11,
       dpi = 600,
       bg = "white",
       path= "Z:/nonwear_detection/preprint_figures/results/fig02")
```

### Descriptive statistics for weekday vs weekend difference
```{r}
daytype_summary <- wknd_wkday %>%
  group_by(day_type) %>%
  summarise(mean = mean(perc_length),
            sd = sd(perc_length),
            min = min(perc_length),
            max = max(perc_length),
            median = median(perc_length),
            range = max-min) %>%
  ungroup()
```

